---
title: "Poly Divr Simulations"
author: "April Peterson"
date: "November 13, 2017"
output: html_document
---

Rmd file for walking through the logic of how I'm setting up my Polymorphism and Divergence simulations.

```{r, setup}
library(plyr)
library(dplyr)
library(ggplot2)

setwd("C:/Users/alpeterson7/Documents/MLH1repo")
load(file="MLH1_data_setup.RData")

PnD_female <- MLH1_data[MLH1_data$sex == "female",]
PnD_male <- MLH1_data[MLH1_data$sex == "male",]

Dom_f <- PnD_female[( PnD_female$strain == "WSB" | PnD_female$strain == "G" |PnD_female$strain == "LEWES"),]
Musc_f <- PnD_female[( PnD_female$strain == "PWD" | PnD_female$strain == "MSM"),]

Dom_m <- PnD_male[( PnD_male$strain == "WSB" | PnD_male$strain == "G" |PnD_male$strain == "LEWES"),]
Musc_m <- PnD_male[( PnD_male$strain == "PWD" | PnD_male$strain == "MSM"),]
Cast_m <- PnD_male[( PnD_male$strain == "CAST" | PnD_male$strain == "HMI"),]

```

Why do these P and D simulations?
To test if the ratios for females and males are different (than a random sample).

Polymorphism measures are SE.
```{r, Polymorphism Measures}
F_poly_table <- ddply(PnD_female, c("strain"), summarise,
                     Nmice = length(unique(mouse)),
                     Ncells  = length(adj_nMLH1.foci),
                     mean_co = format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3),
                     var = format(round(   var(nMLH1.foci),3), nsmall=3),
                     sd   = round(sd(adj_nMLH1.foci), 3),
                     se   = round(sd / sqrt(Ncells), 3),
                     cV = round( (as.numeric(sd) / as.numeric(mean_co) ),3)
)
DomF_poly_var <- var(F_poly_table$mean_co[1:3])
DomF_poly_sd <- sd(F_poly_table$mean_co[1:3])
DomF_poly_se <- (DomF_poly_sd / sqrt(3) )

MuscF_poly_var <- var(F_poly_table$mean_co[2:5])
MuscF_poly_sd <- sd(F_poly_table$mean_co[2:5])
MuscF_poly_se <- (MuscF_poly_sd / sqrt(2))


M_poly_table <- ddply(PnD_male, c("strain"), summarise,
                      Nmice = length(unique(mouse)),
                      Ncells  = length(adj_nMLH1.foci),
                      mean_co = format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3),
                      var = format(round(   var(nMLH1.foci),3), nsmall=3),
                      sd   = round(sd(adj_nMLH1.foci), 3),
                      se   = round(sd / sqrt(Ncells), 3),
                      cV = round( (as.numeric(sd) / as.numeric(mean_co) ),3)
)

DomM_poly_var <- var(M_poly_table$mean_co[1:3])
DomM_poly_sd <- sd(M_poly_table$mean_co[1:3])
DomM_poly_se <- (DomM_poly_sd / sqrt(3))

MuscM_poly_var <- var(M_poly_table$mean_co[2:5])
MuscM_poly_sd <- sd(M_poly_table$mean_co[2:5])
MuscM_poly_se <- (MuscM_poly_sd / sqrt(2) )

CastM_poly_var <- var(M_poly_table$mean_co[6:7])
CastM_poly_sd <- sd(M_poly_table$mean_co[6:7])
CastM_poly_se <- (CastM_poly_sd / sqrt(2) )
```
Above are the Polymorphism measures.

Divergence measures are paired, there will be mulitple divergent measures per strain
```{r, Divergence Measures}
#calq subsp wide averages
DomF_D_mean <- mean(as.numeric(F_poly_table$mean_co[1:3]) )
MuscF_D_mean <- mean(as.numeric(F_poly_table$mean_co[2:5]) )

DomM_D_mean <- mean(as.numeric(M_poly_table$mean_co[1:3]) )
MuscM_D_mean <- mean(as.numeric(M_poly_table$mean_co[2:5]) )
CastM_D_mean <- mean(as.numeric(M_poly_table$mean_co[6:7]) )

# divergence can only be calq in pairwise manner
# Musc - Dom, Musc - spretus(spic), Musc - caroli,
F_D_HMdm_se <- ( sd(c(DomF_D_mean, MuscF_D_mean)) / sqrt(2) )

M_D_HMdm_se <- ( sd(c(DomM_D_mean, MuscM_D_mean) ) / sqrt(2) )
M_D_HMdc_se <- ( sd(c(DomM_D_mean, CastM_D_mean) ) / sqrt(2) )
M_D_HMmc_se <- ( sd(c(CastM_D_mean, MuscM_D_mean) ) / sqrt(2) )

```

Now start simulations for making random samples from the database.
start sampling througheach type of mouse
```{r, simulations}

#3 calqs of SE from random samples
n3 <-  replicate(3, (sd(sample(Dom_f$nMLH1.foci, 30) ) / sqrt(30) ) )
#se of the simulated se's       
se_per <- sd( replicate(3, (sd(sample(Dom_f$nMLH1.foci, 30) ) / sqrt(30) ) ) ) / sqrt(3)

#100 simulated se's / P values for fDom
per100se <- replicate(1000,
                      sd( replicate(3, (sd(sample(Dom_f$nMLH1.foci, 30) ) / sqrt(30) ) ) ) / sqrt(3)
)
#this will show a hist of the simulated P values, with R line for the real
hist(per100se, xlim = range(0,.5) )
abline(v=DomM_poly_se,col="red")

```

#### Comparisons for simulations.

Comparing the histograms of simulated Polymorphism values by the size,



```{r, poly simulations}
#3 calqs of SE from random samples
male_musc <-  replicate(3, (sd(sample(Musc_m$nMLH1.foci, 30) ) / sqrt(30) ) )
#se of the simulated se's       
se_per <- sd( replicate(3, (sd(sample(Musc_m$nMLH1.foci, 30) ) / sqrt(30) ) ) ) / sqrt(3)


#100 simulated se's / P values for fDom
male_musc_ss30 <- replicate(1000,
        sd( replicate(3, (sd(sample(Musc_m$nMLH1.foci, 30) ) / sqrt(30) ) ) ) / sqrt(3)
)
male_musc_ss15 <- replicate(1000,
        sd( replicate(3, (sd(sample(Musc_m$nMLH1.foci, 15) ) / sqrt(15) ) ) ) / sqrt(3)
)
male_musc_ss10 <- replicate(1000,
        sd( replicate(3, (sd(sample(Musc_m$nMLH1.foci, 10) ) / sqrt(10) ) ) ) / sqrt(3)
)

#this will show a hist of the simulated P values, with R line for the real


par(mfrow=c(1, 3))
hist(male_musc_ss30, xlim = range(0,3) )
hist(male_musc_ss15, xlim = range(0,3) )
hist(male_musc_ss10, xlim = range(0,3) )
abline(v=MuscM_poly_se,col="red")

Dom_m <- PnD_male[( PnD_male$strain == "WSB" | PnD_male$strain == "G" |PnD_male$strain == "LEWES"),]
Musc_m <- PnD_male[( PnD_male$strain == "PWD" | PnD_male$strain == "MSM"),]
Cast_m <- PnD_male[( PnD_male$strain == "CAST" | PnD_male$strain == "HMI"),]
```



Compare the effects of drawing MLH1 from specific strains, instead of pooled values.

```{r}
# sample from strain
male_musc_ss10 <- replicate(1000, c(
  sample( PnD_male$nMLH1.foci[PnD_male$strain == "PWD"], 10),
  sample( PnD_male$nMLH1.foci[PnD_male$strain == "MSM"], 10) ) 
)

hist(male_pwd_ss10, xlim = range(15,40) )

male_dom_ss10 <- replicate(1000, c(
  sample( PnD_male$nMLH1.foci[PnD_male$strain == "G"], 10),
  sample( PnD_male$nMLH1.foci[PnD_male$strain == "LEWES"], 10),
  sample( PnD_male$nMLH1.foci[PnD_male$strain == "WSB"], 10)
  ) 
)

#b <- hist(male_dom_ss10, xlim = range(10,40) )
#b <- abline( v=mean( PnD_male$nMLH1.foci[(PnD_male$strain == "WSB")] #), col="red")

 # mean(PnD_male$nMLH1.foci[(PnD_male$strain == "G")] ),
#  mean(PnD_male$nMLH1.foci[(PnD_male$strain == "LEWES")] ) ), col="red")
#PnD_male[( PnD_male$strain == "WSB" | PnD_male$strain == "G" |PnD_male$strain == "LEWES")


male_msm_ss10 <- sample( PnD_male$nMLH1.foci[PnD_male$strain == "MSM"], 10)

male_musc_ss15 <- replicate(1000,
        sd( replicate(3, (sd(sample(Musc_m$nMLH1.foci, 15) ) / sqrt(15) ) ) ) / sqrt(3)
)
male_musc_ss10 <- replicate(1000,
        sd( replicate(3, (sd(sample(Musc_m$nMLH1.foci, 10) ) / sqrt(10) ) ) ) / sqrt(3)
)

```
```{r, divergence simulations}

#simulating divergence --- take the 3 strain sampes from the polymorphism sims
# calq se / var from a constant measure of divergence



#take a list of the polymorphis measures -- pair cbind/rbind with simulations for D, plot the ratios


```

