---
title: "Chapter 2 Figures"
author: "April Peterson"
date: "December 16, 2019"
output: html_document
---

```{r setup, echo=FALSE, warning=FALSE, include=FALSE}
#load data ect
library(knitr)
library(ggplot2)
#library(pwr)
library(plyr)
library(dplyr)
library(lattice)
library(dplyr)
library("cowplot")
library(raster)#for cV

#setwd("~./MLH1repo/")
setwd("C:/Users/alpeterson7/Documents/MLH1repo/")
#setwd("C:/Users/April/Desktop/MLH1repo")
#load(file="data/MLH1/MLH1_data_setup.RData")

#load main data file
load(file="data/MLH1/MLH1_data_setup_12.3.19.RData")#removed exclude mice

source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
```


```{r strain.color, echo=FALSE}

colors_of_strains <- c('WSB'= "#56B4E9",'G' ='cadetblue','LEW'= 'lightblue','PERC'= 'blue',
                       #blues
                       
                       'PWD'= 'red2',
                       'SKIVE'= 'coral1','KAZ'= 'coral4', 'TOM'= 'indianred1','AST' = 'red4',
                       'CZECH'= 'indianred4',
                       
                       'MSM'= 'deeppink', 'MOLF'= 'hotpink', #mol are pinks, bc they are fancy mice
                       
                    'CAST'='green','HMI'= 'forestgreen', #greens
                    
                    'SPRET'='gold','SPIC'= 'goldenrod1', #browns? color of mounds?
                    'CAROLI'='gold3',
                    
                    "F1"='grey', "other" = 'grey'
                    )

```


```{r setting.data, echo=FALSE, warning=FALSE}
#remove F1
AP_mouse_table.HQ <- AP_mouse_table.HQ[(AP_mouse_table.HQ$strain != "F1"),]

#remove outgroups
AP_mouse_table.HQ <- AP_mouse_table.HQ[(AP_mouse_table.HQ$strain != "SPRET"),]
AP_mouse_table.HQ <- AP_mouse_table.HQ[(AP_mouse_table.HQ$strain != "SPIC"),]
AP_mouse_table.HQ <- AP_mouse_table.HQ[(AP_mouse_table.HQ$strain != "CAROLI"),]

#quality control - remove outlier mice
AP_mouse_table.HQ <- AP_mouse_table.HQ[(AP_mouse_table.HQ$mouse != "18nov17_WSB_f5"),] #this mouse only had 1 cell

#remove male only data

#I need cell level data 
uni.mice <- unique(AP_mouse_table.HQ$mouse)
fig.data<- MLH1_data[ MLH1_data$mouse %in% uni.mice, ]

#take all mouse levels from AP.table,
#subset MLH1 data with those mice
```

```{r strain.table.v2, echo=FALSE, warning=FALSE}

#there are still quality 5 cells in this dataframe
fig.data <- subset(fig.data, quality < 5)


#step-wise, 1)mouse-level, 2)averages of mouse-levels


#Pooling all cells for a category
strain.table.HQ <- ddply(fig.data, c("strain", "sex"), summarise,
                        Nmice =length(unique(mouse)),
                        Ncells  =length(nMLH1.foci),
                        mean_co =as.numeric(format(round(mean(nMLH1.foci), 3 ), nsmall=3) ),
                        cV = cv(nMLH1.foci, na.rm = TRUE),
                        var = as.numeric(format(round(   var(nMLH1.foci, na.rm = TRUE),3), nsmall=3)), 
                        sd   = round(sd(nMLH1.foci, na.rm = TRUE), 3),
                       #Standard error (mice vs cells)
                        se  = round(sd / sqrt(Ncells), 3)
    )


#cv, sd and se are all NA / empty
#this removes MOLF female
#strain.table.HQ <- strain.table.HQ[!(is.na(strain.table.HQ$cV) | strain.table.HQ$cV== "" ), ]

#remove outgroups
#remove F1
strain.table.HQ <- strain.table.HQ[(strain.table.HQ$strain != "F1"),]

#remove outgroups
strain.table.HQ <- strain.table.HQ[(strain.table.HQ$strain != "SPRET"),]
strain.table.HQ <- strain.table.HQ[(strain.table.HQ$strain != "SPIC"),]
strain.table.HQ <- strain.table.HQ[(strain.table.HQ$strain != "CAROLI"),]

#REMOVE ALL STRAINS WITHOUT FEMALE DATA
strain.table.HQ <- strain.table.HQ[(strain.table.HQ$strain != "CAST"),]
strain.table.HQ <- strain.table.HQ[(strain.table.HQ$strain != "HMI"),]

strain.table.HQ <- strain.table.HQ[(strain.table.HQ$strain != "PERC"),]
strain.table.HQ <- strain.table.HQ[(strain.table.HQ$strain != "TOM"),]
strain.table.HQ <- strain.table.HQ[(strain.table.HQ$strain != "AST"),]
strain.table.HQ <- strain.table.HQ[(strain.table.HQ$strain != "CZECH"),]

strain.table.HQ <- strain.table.HQ[(strain.table.HQ$strain != "HMI"),]
strain.table.HQ <- strain.table.HQ[(strain.table.HQ$strain != "CAST"),]

```


```{r reorder.strain.v2, echo=FALSE}

#re-order strains
#Musc
#PWD, SKIVE, KAZ, CZECH, TOM, AST

#MOl
#MSM, MOLF
strain.table.HQ$strain<- factor(strain.table.HQ$strain, ordered = TRUE, levels =c( "WSB", "G", "LEW", "PERC",
                                                                   "PWD", "SKIVE", "KAZ", "TOM", "AST","CZECH",
                                                                  "MSM", "MOLF",
                                                                    "CAST", "HMI",
                                                                   "SPRET", "SPIC", "CAROLI", "F1", "other") )
```




# Option 1, strain averages, sexes matched, cells pooled

learning that geom_point(position = position_doge(x))
 will shift the two shaped points closer together.  But this isn't the same as the x axis categories being shifted closer together. 
 
So trying to learn how to squish the x axis together
 scale_x_discrete(expand=c(0, 2) ) squishes all the tick marks, but makes a bunch of big space at the boarders

The parameters in expand() -- are connected with the data range. -- expand is used for making padding around the data.  So this might not be the exactly what I am looking for.
It might be that the best way is to make a stand in continous x axis scale for nudging the point positions.


Removing legends, selectively

has to do with which variables are defined within aes(), not sure how to define shape but not color within aes

the color will now be by sex... I could do this in Illustrator


```{r main.plot.code, echo=FALSE, warning=FALSE}
#use the MLH1 / cell level data to calq the error bars
#strain table ggplot

Fig1_Bv2_strain <- ggplot(strain.table.HQ, aes(y=mean_co, x=strain,shape = sex, color=strain),   show.legend = FALSE)+

    geom_point(aes(shape = sex),  size=3.7, position=position_dodge(.29))+
  
   geom_errorbar(aes(ymin=mean_co-(se*2), ymax=mean_co+(se*2) ), width=.3, position = position_dodge(.29), size=1.001)+
   
  scale_color_manual(values=colors_of_strains)+ 
 
  ylab("Number of Crossovers per Cell")+
  scale_y_continuous(breaks=seq(20,32,2) )+
  
  scale_x_discrete(expand=c(.08, 0) )+  
  #scale_x_continuous( breaks=c(1,2), limits=c(0,3), expand=c(0,0))+
  
  theme_classic()+
  theme(legend.position="none",     #legend.position="bottom",legend.box = "horizontal",
        axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank() )+
 #remove all legend for now....
   
  #  panel.grid.major = element_blank(), panel.grid.minor = #element_blank(),  axis.line=element_line(colour = "black"),
 # panel.background = element_blank()) +
  
#remove y labfor male 
  #Dom
  annotate("segment", x=.75, xend=3.6, y=21, yend=21, colour="black", size=1)+
  #Musc
  annotate("segment", x=3.9, xend=6.15, y=21, yend=21, colour="black", size=1)+
  #mol
  annotate("segment", x=6.5, xend=8.1, y=21, yend=21, colour="black", size=1)+
  #label
  annotate("text", x = c(2.2, 4.5, 7), #list of all x points 
           y = c(20.2,20.2,20.2),
   label = c("M. m. domesticus","M. m. musculus", "M. m. molossinus"),
           fontface = 'italic', hjust = 0.5, size=3.5)

#this can pull just the lengend object
#grobs <- ggplotGrob(Fig1_B_strain.male)$grobs
#legend <- grobs[[which(sapply(grobs, function(x) x$name) == "guide-box")]]
```


```{r Strain.fig.merge.v2, echo=FALSE}
Fig1_Bv2_strain
#save a tif or png file
#ggsave("~./MLH1repo/doc/figureFiles/MLH1Fig1.png")
```


Error bars are 2 SE, 

S.E. = sd( all cells) / n.cells


# FINAL DRAFT

put together 3 pieces; plot, cell1, cell2

 todo  

   1. finish and import final ai edited polt,  

   2. add cell images with cowplot to bottom


```{r load.plots, echo=FALSE}
library(png)
library(ggpubr)

F.first.png <- readPNG('C:/Users/alpeterson7/Documents/MLH1repo/doc/figureFiles/MLH1_F.first.plot_v4.png')
M.first.png <- readPNG('C:/Users/alpeterson7/Documents/MLH1repo/doc/figureFiles/MLH1_M.first.plot_v4.png')

empty.DF <- data.frame(col1 = c(1,3), col2 = c(1,3) )

F.first <- ggplot(empty.DF, aes(x = col1, y = col2)) +
  background_image(F.first.png) + theme(axis.title.x=element_blank(),
                                  axis.text.x=element_blank(),axis.ticks.x=element_blank(),axis.title.y=element_blank(),
                                  axis.text.y=element_blank(),axis.ticks.y=element_blank() )

M.first <- ggplot(empty.DF, aes(x = col1, y = col2)) +
  background_image(M.first.png) + theme(axis.title.x=element_blank(),
                                  axis.text.x=element_blank(),axis.ticks.x=element_blank(),axis.title.y=element_blank(),
                                  axis.text.y=element_blank(),axis.ticks.y=element_blank() )




```


```{r load.cells, echo=FALSE}

best.oo <- readPNG('C:/Users/alpeterson7/Documents/MLH1repo/doc/figureFiles/cell.images/Best_female_4apr16_17mar16_G_f3_sp1_26_rev.png')
best.spermato <- readPNG('C:/Users/alpeterson7/Documents/MLH1repo/doc/figureFiles/cell.images/Best_male_9mar16_22feb16_G_m1_sp1_15.2_rev.png')
  
#make empty DF for ggplot object
empty.DF <- data.frame(col1 = c(1,3), col2 = c(1,3) )
#make an empty ggplot thing, then background ontop
spermatocyte.auto <- ggplot(empty.DF, aes(x = col1, y = col2)) +
  background_image(best.spermato) + theme(axis.title.x=element_blank(),
                                  axis.text.x=element_blank(),axis.ticks.x=element_blank(),axis.title.y=element_blank(),
                                  axis.text.y=element_blank(),axis.ticks.y=element_blank() )


oocyte.auto <- ggplot(empty.DF, aes(x = col1, y = col2)) +
  background_image(best.oo) + theme(axis.title.x=element_blank(),
                                  axis.text.x=element_blank(),axis.ticks.x=element_blank(),axis.title.y=element_blank(),
                                  axis.text.y=element_blank(),axis.ticks.y=element_blank() )


```


# all png loaded in


```{r cow.plot, echo=FALSE}
library("cowplot")

F.first.fig  <- ggdraw() +
  draw_plot(F.first, x = 0, y = .42, width=1, height=.5) +
  
  draw_plot(spermatocyte.auto, x=.51, y = 0, width=.31, height=.39) +
  
  draw_plot(oocyte.auto, x=0.01, y = 0, width = .31, height = .39) +
  
  draw_plot_label(label = c("A", "B", "C"), size = 15,
                  x = c(0, 0, .5), y = c(1, .43, 0.43) )

F.first.fig

#ggsave("~./WildMiceMadison/results/Figure1_32119.png")
```

Loaded slightly nudged points-plot as png. This looks bad, don't do this. Switch to Illustrator.


# redo


```{r redo.plot, echo=FALSE}
redo  <- ggdraw() +
  
  draw_plot(Fig1_Bv2_strain, x = 0, y = .42, width=1, height=.49) +
  
  draw_plot(spermatocyte.auto, x=.55, y = 0, width=.4, height=.41) +
  
  draw_plot(oocyte.auto, x=0.05, y = 0, width = .4, height = .41) +
  
  draw_plot_label(label = c("A", "B", "C"), size = 15,
                  x = c(0, 0, .5), y = c(1, .43, 0.43) )

redo
```


All pieces are 'made' within R. But I can't figure out how to reduce the x axis white space between points (strains). The message boards seem to indicate that using a continous_x scale might be the only way to get the layout I want.

# Option vertical facets



```{r vertical.facet, echo=FALSE, fig.height= 13}

Vert.facet <- ggplot(strain.table.HQ, aes(y=mean_co, x=strain, color=strain, shape = sex))+
  #facet_grid(~sex, scales = "free", space="free")+ 
  ylab("Number of Crossovers per Cell")+
  geom_point(aes(shape = sex), size=3, position=position_dodge(.5))+

   facet_wrap(~sex, ncol = 1)+ #, scales = "free", space="free"
  
   geom_errorbar(aes(ymin=mean_co-(se*2), ymax=mean_co+(se*2) ), width=.3, position = position_dodge(.5), size=1.001)+
  
  ylim(c(17.5, 33))+
 scale_color_manual(values=colors_of_strains)+ 
  scale_y_continuous(breaks=seq(20,32,2) )+
  
  #legend.position="none",
  theme(legend.position="bottom",legend.box = "horizontal",
  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(), axis.line=element_line(colour = "black"),
  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())+
  
#remove y labfor male 
  #Dom
  annotate("segment", x=.75, xend=3.6, y=21, yend=21, colour="black", size=1)+
  
  #Musc
  annotate("segment", x=3.9, xend=6.15, y=21, yend=21, colour="black", size=1)+
  #mol
  annotate("segment", x=6.5, xend=8.1, y=21, yend=21, colour="black", size=1)
 
   #label -- segments applied to both facets, labels not?
#label annotation not working, add in illustrator  
annotate("text", x = c(2.2, 4.5, 7,
                         2.2, 4.5, 7), #list of all x points 
           y = c(20.2,20.2,20.2,
                 20.2,20.2,20.2),
   label = c("M. m. domesticus","M. m. musculus", "M. m. molossinus",
             "M. m. domesticus","M. m. musculus", "M. m. molossinus"),
           fontface = 'italic', hjust = 0.5, size=3.5)

  Vert.facet
  
```


```{r new.save, echo=FALSE}
#Fig1_Bv2_strain
#save a tif or png file
#ggsave("~./MLH1repo/doc/figureFiles/MLH1Fig1.png")
```





## Option 2, mouse averages, more strains


```{r strain.table, echo=FALSE, warning=FALSE}
#fig.data -- (only mice from better list)

#there are still quality 5 cells in this dataframe
fig.data <- subset(fig.data, quality < 5)

mouse.table.HQ.backup <- ddply(fig.data, c("mouse"), summarise,
                        Ncells  = length(nMLH1.foci),
                        mean_co = as.numeric(format(round(  mean(nMLH1.foci), 3 ), nsmall=3) ),
                        cV = cv(nMLH1.foci, na.rm = TRUE),
                        var = as.numeric(format(round(   var(nMLH1.foci, na.rm = TRUE),3), nsmall=3)), #this turns to chr
                        sd   = round(sd(nMLH1.foci, na.rm = TRUE), 3),
                        se   = round(sd / sqrt(Ncells), 3) #mean is pooled cells, so sd should be pooled
    )
#need to add strain and sex
mouse.table.HQ.backup <- add_strain(mouse.table.HQ.backup)
mouse.table.HQ.backup <- add_sex(mouse.table.HQ.backup)


strain.table.from.mouse.HQ.backup <- ddply(mouse.table.HQ.backup, c("strain","sex"), summarise,
                  Nmice  = length(mouse),
                  mean_co.of.mice = as.numeric(format(round(  mean(mean_co), 3 ), nsmall=3) ),
                  cV.of.mice = cv(mean_co, na.rm = TRUE),
                   var.of.mice = as.numeric(format(round(   var(mean_co, na.rm = TRUE),3), nsmall=3)), #this turns to chr
                  sd.of.mice   = round(sd(mean_co, na.rm = TRUE), 3),
                  se.of.mice   = round(sd.of.mice / sqrt(Nmice), 3) #mean is pooled cells, so sd should be pooled
    )



#use the MLH1 / cell level data to calq the error bars
strain.table.HQ.backup <- ddply(fig.data, c("strain", "sex"), summarise,
                        Nmice = length(unique(mouse)),
                        Ncells  = length(nMLH1.foci),
                        mean_co = as.numeric(format(round(  mean(nMLH1.foci), 3 ), nsmall=3) ),
                        cV = cv(nMLH1.foci, na.rm = TRUE),
                        var = as.numeric(format(round(   var(nMLH1.foci, na.rm = TRUE),3), nsmall=3)), #this turns to chr
                        sd   = round(sd(nMLH1.foci, na.rm = TRUE), 3),
                        se   = round(sd / sqrt(Ncells), 3) #mean is pooled cells, so sd should be pooled
    )


#cv, sd and se are all NA / empty
#strain.table.HQ.backup <- strain.table.HQ.backup[!(is.na(strain.table.HQ.backup$cV) | strain.table.HQ.backup$cV== "" ), ]

#this removes MOLF female

#remove outgroups
#remove F1
strain.table.HQ.backup <- strain.table.HQ.backup[(strain.table.HQ.backup$strain != "F1"),]

#remove outgroups
strain.table.HQ.backup <- strain.table.HQ.backup[(strain.table.HQ.backup$strain != "SPRET"),]
strain.table.HQ.backup <- strain.table.HQ.backup[(strain.table.HQ.backup$strain != "SPIC"),]
strain.table.HQ.backup <- strain.table.HQ.backup[(strain.table.HQ.backup$strain != "CAROLI"),]

strain.table.HQ.backup <- strain.table.HQ.backup[(strain.table.HQ.backup$strain != "CAST"),]
strain.table.HQ.backup <- strain.table.HQ.backup[(strain.table.HQ.backup$strain != "HMI"),]

#Keep more of the strains
#strain.table.HQ <- strain.table.HQ[(strain.table.HQ$strain != "PERC"),]
#strain.table.HQ <- strain.table.HQ[(strain.table.HQ$strain != "TOM"),]
#strain.table.HQ <- strain.table.HQ[(strain.table.HQ$strain != "AST"),]
#strain.table.HQ <- strain.table.HQ[(strain.table.HQ$strain != "CZECH"),]

#strain.table.HQ <- strain.table.HQ[(strain.table.HQ$strain != "HMI"),]
#strain.table.HQ <- strain.table.HQ[(strain.table.HQ$strain != "CAST"),]

```


```{r reorder.strain, echo=FALSE}

#re-order strains
#Musc
#PWD, SKIVE, KAZ, CZECH, TOM, AST

#MOl
#MSM, MOLF
strain.table.from.mouse.HQ.backup$strain<- factor(strain.table.from.mouse.HQ.backup$strain, ordered = TRUE, levels =c( "WSB", "G", "LEW", "PERC",
                           "PWD", "SKIVE", "KAZ", "TOM", "AST","CZECH",
                                                    "MSM", "MOLF",
                                          "CAST", "HMI",
                             "SPRET", "SPIC", "CAROLI", "F1", "other") )
```


```{r strain.table.plot, echo=FALSE, warning=FALSE}
#use the MLH1 / cell level data to calq the error bars
#strain table ggplot


#REMOVING FACET ~SEX 
Fig1_B_strain <- ggplot(strain.table.from.mouse.HQ.backup, aes(y=mean_co.of.mice, x=strain, color=strain, shape=sex))+ 
  ylab("Average MLH1 Foci per Cell")+

  geom_point(aes(shape = sex), size=3, position=position_dodge(.5))+
  geom_errorbar(aes(ymin=mean_co.of.mice-(se.of.mice*2), ymax=mean_co.of.mice+(se.of.mice*2) ), width=.3, position = position_dodge(.5), size=1.001)+
  
  
  ylim(c(18, 35))+
 scale_color_manual(values=colors_of_strains)+ 
  theme(legend.position="none",
  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(), axis.line=element_line(colour = "black"),
  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())+
#remove y labfor male 
  #segments
  annotate("segment", x=.75, xend=4.5, y=19.5, yend=19.5, colour="black", size=1)+
  #Musc
  annotate("segment", x= 5.5, xend=10.2, y=19.5, yend=19.5, colour="black", size=1)+
  #mol
  annotate("segment", x=11, xend=12.25, y=19.5, yend=19.5, colour="black", size=1)+
  #label
  annotate("text", x = c(3,8.5,10), #list of all x points 
           y = c(18.5,18.5,33),
           
   label = c("M. m. domesticus","M. m. musculus", "M.m.molossinus"),
           fontface = 'italic', hjust = 0.5, size=3.5)

#this can pull just the lengend object
#grobs <- ggplotGrob(Fig1_B_strain.male)$grobs
#legend <- grobs[[which(sapply(grobs, function(x) x$name) == "guide-box")]]

# 2
Fig1_B_strain.female <- ggplot(strain.table.HQ.backup[strain.table.HQ.backup$sex == "female",], aes(y=mean_co, x=strain, color=strain))+facet_grid(~sex, scales = "free", space="free")+ ylab("Average MLH1 Foci per Cell")+
  geom_point(size=3)+
  ylim(c(18, 35))+
  geom_errorbar(aes(ymin=mean_co-se, ymax=mean_co+se), width=.3) +
  scale_color_manual(values=colors_of_strains)+ 
  theme(legend.position="none", panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(), axis.line = element_line(colour = "black"),
  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())+
  #segments
  annotate("segment", x=.75, xend=3.5,y=19.5, yend=19.5, colour="black", size=1)+
  #Musc
  annotate("segment", x=4, xend=6.5, y=19.5, yend=19.5, colour="black", size=1)+
#mol
 annotate("segment", x=7, xend=8.5, y=19.5, yend=19.5, colour="black", size=1)+
    #Cast
#  annotate("segment", x=4, xend=2, y=18, yend=18, colour="black", size=1)
#Labels
 annotate("text", x = c(2.25,5.6,6.5), #list of all x points 
           y = c(18.5,18.5,33),
           label = c("M. m. domesticus","M. m. musculus", "M.m.molossinus"),
           fontface = 'italic', hjust = 0.5, size=3.5)
 

```


```{r Strain.fig.merge, echo=FALSE}
#fig1_B_v1  <- ggdraw() +
#  draw_plot(Fig1_B_strain.female, x=0, y = 0, width=.45, height=.97) +
#  draw_plot(Fig1_B_strain.male, x=.5, y=0, width=.45, height=.97) +
  
#  draw_plot_label(label = c("A", "B"), size = 15,
#                  x = c(0, .5), y = c(1, 1) )


Fig1_B_strain

```



This is currently backup version.

I removed the cast and outgroup measures. I left off the color legend, but I think I could add one to the bottom of the figure. These show the strain averages, but I can try adding the mouse averages, if you think having mulitple points would show the data better.

## Option 2.B strain averages, mouse observations

```{r strain.table.plot.mouse.level, echo=FALSE, warning=FALSE}
#use the MLH1 / cell level data to calq the error bars
#strain table ggplot

Fig1_C_strain.male <- ggplot(strain.table.from.mouse.HQ.backup[strain.table.from.mouse.HQ.backup$sex == "male",], aes(y=mean_co.of.mice, x=strain, color=strain))+facet_grid(~sex, scales = "free", space="free")+ 
  ylab("")+
  geom_point(size=3)+
  geom_errorbar(aes(ymin=mean_co.of.mice-(se.of.mice*2) , ymax=mean_co.of.mice+(se.of.mice*2) ), width=.3)+
  ylim(c(18, 35))+
 scale_color_manual(values=colors_of_strains)+ 
  theme(legend.position="none",
  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(), axis.line=element_line(colour = "black"),
  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())+
#remove y labfor male 
  #segments
  annotate("segment", x=.75, xend=4.5, y=19.5, yend=19.5, colour="black", size=1)+
  #Musc
  annotate("segment", x= 5.5, xend=10.2, y=19.5, yend=19.5, colour="black", size=1)+
  #mol
  annotate("segment", x=11, xend=12.25, y=19.5, yend=19.5, colour="black", size=1)+
  #label
  annotate("text", x = c(3,8.5,10), #list of all x points 
           y = c(18.5,18.5,33),
           
   label = c("M. m. domesticus","M. m. musculus", "M.m.molossinus"),
           fontface = 'italic', hjust = 0.5, size=3.5)

#this can pull just the lengend object
#grobs <- ggplotGrob(Fig1_B_strain.male)$grobs
#legend <- grobs[[which(sapply(grobs, function(x) x$name) == "guide-box")]]

# 2
Fig1_C_strain.female <- ggplot(strain.table.from.mouse.HQ.backup[strain.table.from.mouse.HQ.backup$sex == "female",], aes(y=mean_co.of.mice, x=strain, color=strain))+facet_grid(~sex, scales = "free", space="free")+ ylab("Average MLH1 Foci per Cell")+
  geom_point(size=3)+
  ylim(c(18, 35))+
  geom_errorbar(aes(ymin=mean_co.of.mice-(se.of.mice*2), ymax=mean_co.of.mice+(se.of.mice*2)  ), width=.3) +
  scale_color_manual(values=colors_of_strains)+ 
  theme(legend.position="none", panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(), axis.line = element_line(colour = "black"),
  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())+
  #segments
  annotate("segment", x=.75, xend=3.5,y=19.5, yend=19.5, colour="black", size=1)+
  #Musc
  annotate("segment", x=4, xend=6.5, y=19.5, yend=19.5, colour="black", size=1)+
#mol
 annotate("segment", x=7, xend=8.5, y=19.5, yend=19.5, colour="black", size=1)+
    #Cast
#  annotate("segment", x=4, xend=2, y=18, yend=18, colour="black", size=1)
#Labels
 annotate("text", x = c(2.25,5.6,6.5), #list of all x points 
           y = c(18.5,18.5,33),
           label = c("M. m. domesticus","M. m. musculus", "M.m.molossinus"),
           fontface = 'italic', hjust = 0.5, size=3.5)
 

```


```{r Strain.fig.merge.mouse.level, echo=FALSE}
fig1_C_v1  <- ggdraw() +
  draw_plot(Fig1_C_strain.female, x=0, y = 0, width=.5, height=.97) +
  draw_plot(Fig1_C_strain.male, x=.45, y=0, width=.5, height=.97) +
  
  draw_plot_label(label = c("A", "B"), size = 15,
                  x = c(0, .5), y = c(1, 1) )
fig1_C_v1
```


Error bars above are 2 S.E. of mouse average COs, 

S.E. = sd( mouse average COs ) / Nmice


MOLF and SKIVE female data point is missing because there is only 1 female, so NAs were generated.
Mol text should be moved back above the points.


## Option 3, mouse boxplots


```{r fig1_scatter_basic, echo=FALSE, warning=FALSE}
#mouse averages
#AP_mouse_table.HQ
#fig.data  cell data


fig.data <- fig.data %>%
  arrange(strain, sex, mouse) %>%
  mutate(Original.Name = factor(Original.Name)) #another category that you want the order to match

# sort your dataframe, by the focal categories
fig.data$mouse <- factor(fig.data$mouse, levels=unique(fig.data$mouse))

fig.data.male <- subset(fig.data, sex== "male")

#male
Fig1_A_mouse.male <- ggplot(fig.data.male[order(fig.data.male$strain), ],
  aes(y=nMLH1.foci, x=as.factor(mouse), color=strain))+  #geom_jitter()+
  geom_boxplot(alpha = 0.6)+ylim(c(12,48))+
  facet_grid(~sex, scales = "free", space="free")+scale_color_manual(values=colors_of_strains)+geom_boxplot(alpha = 0.6)+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),legend.position="none",
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
  #lines -- these might be tricky with the faceting
  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())
  #Dom
#  annotate("segment", x= 10, xend=20.5, y=15, yend=15, colour="black", size=1)+
  #Musc
#  annotate("segment", x= 50, xend=55, y=15, yend=15, colour="black", size=1)+
  #Cast
#  annotate("segment", x=70, xend=80, y=15, yend=15, colour="black", size=1)


#female
#the data is mouse mean 
fig.data.female <- subset(fig.data, sex== "female")

Fig1_A_mouse.female <- ggplot(fig.data.female[order(fig.data.female$strain), ], aes(y=nMLH1.foci, x=as.factor(mouse), color=strain))+ #geom_jitter(width = 0.25)+
  geom_boxplot(alpha = 0.6)+ylim(c(12,48))+
  facet_grid(~sex, scales = "free", space="free")+scale_color_manual(values=colors_of_strains)+
 theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),legend.position="none",
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
  #lines -- these might be tricky with the faceting
 axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())
  #Dom
#  annotate("segment", x= 10, xend=20.5, y=18, yend=18, colour="black", size=1)+
  #Musc
#  annotate("segment", x= 50, xend=55, y=18, yend=18, colour="black", size=1)+
  #Cast
#  annotate("segment", x=70, xend=80, y=18, yend=18, colour="black", size=1)

```


```{r cow.plot.fig1, echo=FALSE}

fig1  <- ggdraw() +
  draw_plot(Fig1_A_mouse.female, x=0, y = 0, width=.48, height=.88) +
  draw_plot(Fig1_A_mouse.male, x=.48, y=0, width=.48, height=.88) +
  draw_plot_label(label = c("A", "B"), size = 15,
                  x = c(0, .5), y = c(1, 1) )

fig1
```

This combination of points and box plots is similar to the final peromyscus figure 1, but the number of individual mice observations makes it very busy.

This might be a good format if fewer mice were used for each strain.


