---
title: "Discussion Draft"
date: ''
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '5'
    toc_float: yes
  pdf_document: default
  word_document:
    toc: yes
    toc_depth: '5'
bibliography: refs_Main_Report.bib
---

```{r format.Notes, include=FALSE}
#table, wite by all of the columns and rows

#over-view of the synthesis of hypothesis testing
#(this is how we're synthsizing / testing / the framework were using to compare the results to the literature)

#constrast on model types
#-evolution
#-models accounting for chromosome based traits

#HetC traits
# 1. typical landscape
# 2. female variance
# 3. chromatin organization

#faster male evolution traits
# 1. Reversed Direction of Heterochiasmy
# 2. correlated gwRR increase and interference strength

#go through each traits and point out which models support each of the traits
```


```{r Model.testing, include=FALSE}

#imortant of testable hypotheses

#model not included in this comparison


```


We performed hypothesis testing of models from the literature and adapted here fit the main results from this study. We choose to examine three modifier models and two functional and cell physiology models (table X). It is a bit challenging to test all of these models together since they were built with difference parameters and built to describe different patterns of recombination variation. The modifier models were built to explain heterochiasmy variation between sexes and the COM functional model was built to describe the interference pattern and difference in the recombination landscape between the sexes. In cases where the models can't be extended to predicting results, we list as 'NA'.


# Reversed heterochiasmy direction

```{r reversed.HetC, echo=FALSE}

#our results of reversed hetC are novel due to the short evolutionary disances across the panel of mouse strains

#spindle selection modle points

```

Female biased heterochiasmy is the most common pattern observed, however male biased heterochiasmy is not especially rare [@brandvain2012scrambling]. Our results of male biased heterochiasmy is novel due to the shorter evolutionary distances between the examined strains.

The direction of heterochiasmy under the **two locus modfier model** is dependent on female specific recombination modifier suppresses or increases recombination, if the modifier is linked to the driving locus, and the meiotic stage (reduction or equatorial division) that the driver acts (see Table 1 of @brandvain2012scrambling).


**strict SAC drives directional selection**,  
**genetic variants which lead to 'unstable chromosome structure will be selected agaisnt**  
**sex differences in SAC strength are connected to conserved differences in gametogensis**  

Under the **spindle based** model, faster male evolution of the genome wide recombination rates is primarily driven the strength of the spindle assembly checkpoint (SAC) acting on the chromosome structure at metaphase alignment. We refer the reader to [@laneKauppi2019] for more details on the SAC pathway.

The transition from metaphase to anaphase is irreversible, so prevention of annueploidy requires that all bivalents be correctly aligned before meiosis can progress (cite). Unstable bivalents detected on the MI spindle, cause the SAC to delay anaphase and allow time for the bivalent to stabilize / connect to the microtubule spindle. If the 'unattached' / unstable bivalents fail to attach, the SAC triggers apoptosis. Genetic variants which increase the number of unstable bivalents, will be selected against in favor for genetic variants causing faster stabilization of bivalents on the MI spindle.  

Multiple lines of evidence indicate that the effectiveness of SAC in spermatogenesis is greater compared to oogenesis (LaneKauppi, cite). The different the detection of achiasmated bivalents (), the rates of annuploidy in mature gametes, and stronger infertility phenotypes in mutants of key SAC genes (cite).

These differences in SAC strength have been connected to conserved features of gametogenesis; centrosome spindle and cell volume (cite).

-- Supporting this stage/cell physiology based mechanism  influencing broad sexual dimorphic phenotypes in recombination landscape. (leading to distinct evolutionary trajectories) **directional selection** in males and relaxed selection in females.

<leaky oocyte SAC linked to acentrosome spindle (multiple aMTOCs) and cell volume, these are features which differ in spermatocyttes (small, centrosome spindle, 2MTOCs)>

<Directional selection could be the result in spermatogenesis if the fitness landscape for 'chromosome structure' is altered via genetic architecture changes or GxE effects. (in oogenesis the leaky SAC has weak selection coefficient)>


# Greater between cell variation


```{r within.mouse.var, include=FALSE}
#review results, other data

#2. Greater between cell variation in females, leaky SAC in oocytes can cause more between cell variation within an individual.

#2a. relaxed selection on chromosome structure in oocyte / leaky SAC, can lead to larger variance in crossover number between cells. The opposite dynamics of the purifying selection for 'strict' SAC in spermatocytes.

#?2b. (leaky SAC could be adaptive/prevent to selfish genetic elements (Lampson))?

```


Our results of greater between cell variation in females, these results are supported by human and mouse (Gruhn, Lynn Kong). [@lynn2002].

These data are also scant, between cell variance isn't often measured especially for both sexes. (given they require single cell observations for both sexes).

While the modifier models parameterize sexual dimorphism in the strength of selection which can shape the variance in a trait these types of models make predictions  for variance between individuals but not within individuals where the genetics are constant. (selection on penetrance).

In the spindle based selection model, the source of between cell variance is attributed to difference strength of checkpoints.
As mentioned above, there is sexual dimorphism in the stringency of the spindle assembly checkpoint (SAC). And these are linked to conserved features of gametogenesis.

The opposite dynamics of the purifying selection for the 'strict'  SAC in spermatocytes is relaxed selection from a 'permissive' SAC. As mentioned above a fundamental sex difference in the metaphase spindle is the presence of centrioles in spermatocytes which anchors and nucleates the microtubulue spindle at the two cell poles. While oocytes lack centrioles which results in a difuse network of microtubules with mulitple microtubule organizing centers (aMTOCs) (Schuh Ellenburg 2007). This or larger cell volumne may lead/be a potnetial source of the leaky SAC in oocytes. Leaky SAC, relaxed selection on the anaphase to metaphase transition could increase the between cell variance for crossover number  


# Typical landscape


```{r typical.landscape.outline, include=FALSE}

# typical landscape is conserved -- sexually dimorphic meiotic trait (Sardell)

# models

#<spindle model -- the telomeric placement of crossovers, minimizes the amount of sister cohesion connecting homologs>
#3. Sexual dimorphism in broad scale recombination landscape,  telomere bias vs uniform recombination landscapes fit with purifying selection and relaxed selection respectively.
#3a. There are sex differences in demand for chromosome cohesion at MI, the trait / variable causing variation in fitness landscape across chromosome structures might be the amount of sister-cohesion connecting homologs.  In oocytes the cell cycle is slower and arrested in some species, this likely selects for chromosome structures that have stability on the spindle. (required to start with more since the cohesion isn't replenished. The optimal chromosome structure might have more sister cohesion, however given the leaky SAC, selection may not be effective.

#3c. Logic of number and placement modulating this area of sister cohesion and chromosome structure (I plan to make a figure to illustrate)
#-For single crossover acrocentric chromosome the sister cohesion distal to the crossover connects homologs. This area reduced when crossover position shifted to telomere (each end of chromosomes). Two crossover acrocentric chromosome the area is distal from the first until the second crossover.

#3d. Metaphase to anaphase transition proceeds by proteolytic decay of sister cohesion on 'arms'.
#- The telomere bias could have an effect of synchronizing the separation of all the bivalents within a cell.

#*caveats: Type II crossovers not accounted for and chromatid interference effects on (multi-crossover chromsomes) could alter the sister cohesion pattern.

```

Our results on the broad scale pattern of crossover placement are supported by the literature. As reviewed in [@sardell_sex_2020], sexual dimorphism in the broad scale recombination landscape is a highly conserved trait.  

The **two locus modifier** and **SACE** model make predictions for sexual dimorphism in the broad scale recombination landscape, for diminishing the effect of drive systems and maintaining larger chromosome blocks respectively.  

The two mechanical models also both predict sexual dimorphism in the recombination landscape for distinct reasons. The COM model (chromosome pairing based model), predicts the sex differences in the positioning of crossovers is due to a combination of the length of the axis and intensity of the rapid prophase movements (RPMs) [@hulten2011_COM, @rubin2020mixing].  


Under the **spindle based model** we hypothesize that the sexual dimorphism in the recombination landscapes hinges on sex differences on the requirements for chromosome cohesion in late meiosis I. The irreversible process of the metaphase to anaphase transition is initiated by the protyltic decay of the sister cohesion connecting homologs (cite). The number and placement of crossovers alter the distribution of sister cohesion and the resulting 'chromosome structure' when bivalents are aligned on the metaphase plate (cite). The male crossover telomere bias could be adaptive in order to synchronize (or minimize) the timing of homologs separating at metaphase I.	Differing lengths of the timing of cell cycle between oogenesis and spermatogenesis may impose the the different selective pressures. Faster spermatogenesis may select for synchronization of the separation homologs or faster separatation at the metaphase to anaphase transition. While in oogenesis, the slower cell cycle and prolonged arrest may require chromosome structures with greater stability on the MI spindle.


# Chromatin Structure

```{r chromatin.organization, include=FALSE}

#supporting results from lit
#chromatin organization

#models
# none of the models make predictions for the axis length (chromatin organization)
# only COM predicts that sexual dimorphism in axis length is due to cell vol

```

Our results from musculus^MSM^ and musculus^PWD^ show that the SC length and chromatin compaction is uncoupled from the direction of heterochiasmy. These results slightly depart from predictions which nominate chromatin organization (axis length) as the primary driver of heterochiasmy [@petkov2007].  While the relationship may hold for more heterochiasmy patterns, but these results indicate that rapid male specific  evolution in gwRR, is through a different trait than chromatin compaction). 

<The results of sexual dimorphism in chromatin structure is supported by the literature, especially for mammals [@lynn2002, @gruhn2013], But there's generally a lack of data for other systems see [@cahoonLibuda2019]).>

Only the COM model predicts sexual dimorphism in chromatin structure will be longer in oocytes due to greater cell volume. This prediction / model might fit broader pattern such as in At where pollen is the larger cell and has longer axis length [@cahoonLibuda2019].  


# Evolution of Interference

```{r evo.interference, include=FALSE}
#review results,
#why unexpected, unexpected pattern is great for identifying most parimonious model / explanation
#(we're gonna spend alot of space on this section)

# more convincing that the pattern is real

#lil background, for interferences 
#assuptions for predictions (most modifer models, )

#models that predict this pattern (CONTROL FOR CO NUMBER -- PRIORITIZE THE WEIGHT /PROB of CO ASSURANCE)

#the pattern is driven in part by short IFD (when short IFD removed in reduced data set -- the low strain mean goes up, but they are still shrter)

#**The rapid evolution of the genome wide rate in our high rec strains is PARTLY due to more enrichment of 2CO which have distinct rec landscape from the 2COs in low rec strains.**

#5. Interference evolution shows a positive correlation with gwRR
#5a. Stronger interference for two crossover chromosomes, increases the amount of sister-cohesion connecting the homologs.
#5b. Under spindle selection model, the fitness landscape of optimal chromosome structure may have changed in the high recombining strains resulting in more 2CO, with larger IFD (sister cohesion area).

```


**we find support that interference strength has evolved in the two groups of male strains. There is a small number of positive correlations between genome wide recombination rate and interference strength in the literature.** 

Interference is a fundamental aspect genome-wide recombination rate the, chromatin compaction and the resulting recombination landscape.

Logically a negative correlation is expected; increasing the number of crossovers across chromosomes would most logically be done by  more densely spacing crossovers along chromosomes and decreasing interference strength. This pattern has empirical support from the most species [@ottoPaysuer2019] and fits well with  the fundamental relationship between the SC area or axis length, the physical upper limit for the number of crossovers. **although it is not particulary strong in our results** We argue that this is due to difference in scale (genome wide vs single chromosome)

Our results support an opposite pattern predicted from the 'logical model', a positive correlation between interference strength and genome wide recombination rates; The within sex comparison of two breeds of cattle with different genome wide recombination rates  [@ma2015_cattle], between lab and wild mice of Peromyscus leucopus, [@peterson2019].

Theoretical models haven't really considered the evolution of interference strength, neither the haploid selection or two locus modifier model can not be applied to evolution of interference strength. While the SACE modifier model does not explicitly model evolution of interference strength we note that a logical outcome of the main prediction of maintaining larger chromosome blocks in males, would be a landscape with stronger interference strength. The COM model predicts that interference and the recombination landscape arises from known oscillatory movements during prophase, it lacks evolutionary based predictions. We propose the spindle based selection model would support the evolution of interference strength in the positive direction via modulation of the amount of sister cohesion connecting homologs (figure).  

Models from Goldstein et al (review in [@ottoPaysuer2019] suggest that if this pattern is widespread interference evolves whenever increased recombination rates evolve. Perhaps a distinguishing feature of models which come to this finding is that the number of crossovers is kept constant. The space across multiple loci (veller?) or between multiple crossovers increases in a positive manner with genome wide recombination rates. Given that the empirical range of crossovers per chromosome is quite small (1-3 [@ottoPaysuer2019, @stapley_variation_2017] and  the obligate crossover rule, the assumption of constraining the number of crossovers per chromosome fits well with empirical data.  


Biological exceptions to interference:  experimentally increasing crossover number through mutants or fusion chromosomes  (Celegans, (plants, fungi with negative interference.


#Future Steps

```{r future.step.plan, echo=FALSE}
#still preliminary
```

- sex-specific and sex average biases. Some cytology models (Celegans) mostly focus on oogenesis  and genetic maps are often from hermaphrodites. (lack of data from one sex). Recognize the value of Chromosome level measures and incorporating chromosome behavior into models.

Broader communication across sub-fields of cell physiology and biophysics for understanding recombination rate evolution. Utilize mitosis systems given the conservation of the molecular pathway. Think of meiosis as a program.

Dark matter of the genome. We can't really exclude that some of the recombination rate variation is due to structural variation (more of less genome). (since everything is mapped to B6). (also maybe there are crossovers in the centromeres and telomeres -- which might escape detected in linkage maps)

what are the rates of sister-sister crossovers?

How does elevated rates of precocious sister centromere separation effect oogenesis?

Are selfish elements affecting the evolution of meiosis -- and thus recombination rates (from a functional standpoint).

Keep generating those testable hypotheses (cite).  

- chromatin interference

#References
