---
title: "DMC1 report"
author: "April Peterson"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---


```{r setup, echo=FALSE, include = FALSE}
library(knitr)
library(ggplot2)
#library(pwr)
library(plyr)
library(lattice)
library(dplyr)
library(raster)
library(lme4)
library(nlme)
library(RLRsim)
library(ggpubr)

setwd("C:/Users/alpeterson7/Documents/MLH1repo/")
#setwd("C:/Users/April/Desktop/MLH1repo")
#load(file="data/MLH1/MLH1_data_setup.RData")

#load main data file

load(file="data/MLH1/MLH1_data_setup_11.12.19.RData")#added batch17
#load(file="data/MLH1/MLH1_data_setup_11.11.19.RData")#deleted DUP image
#load(file="data/MLH1/MLH1_data_setup_9.26.19.RData")#added batch16
#load(file="data/MLH1/MLH1_data_setup_8.29.19.RData") #
#load(file="data/MetaData.RData") #this was replacing most updated data file

#load functions
source("~./MLH1repo/src/CommonFunc_MLH1repo.R")

#loat DMC1 csv
DMC1.DF <- read.csv("~./DMC1/newJuviQuant.csv", header = TRUE)
DMC1.DF$fileName <- DMC1.DF$Original.Name
DMC1.DF <- add_mouse(DMC1.DF)
DMC1.DF <- add_strain(DMC1.DF)
DMC1.DF <- add_subsp(DMC1.DF)
DMC1.DF <- DMC1.DF[DMC1.DF$stage != "P",]

```


#DMC1 Results

ToDo:
  
  - attempt to supplement data with more KAZ DMC1 cell observations.
  - change all the bad variable names




```{r DMC1.plot, echo=FALSE, warning=FALSE}
#think of better color scheme
#make the jitter col thinner or make the stacked dots.
#remember that box-plots add outlier dots.
#change y labels


#https://www.r-bloggers.com/add-p-values-and-significance-levels-to-ggplots/

DMC1.DF$strain<- factor(DMC1.DF$strain, levels =c("G","WSB","KAZ",
              "PWD","MSM"), order=T )
#order the data frame
DMC1.DF <- with(DMC1.DF, DMC1.DF[order(strain),])

#for the test bars, used with STAT_COMPARE
my_comparisons <- list( c("KAZ", "MSM"))#no idea how to adjust


#experimenting with ggboplot
everyday <- ggboxplot(DMC1.DF[DMC1.DF$stage != "P",], x = "strain", y = "total.count",
          color = "stage", palette = "jco",
          add = "jitter", short.panel.labs = FALSE)
  #library gpubr has functions to add in sig/test bars.

DMC1.boxplot <- ggplot(aes(x=strain, y=total.count, color=stage), data=DMC1.DF[DMC1.DF$stage != "P",])+
  geom_boxplot(aes(alpha=.5))+geom_jitter()+ggtitle("DMC1 averages in juvenile mice")
  
  #this will add neat bars, but still need to figure out what groups ect it's actually using
#THE stat_compare function adds a arm thing and pvalue for 
#  stat_compare_means(comparisons = my_comparisons)

#status - legend, expression y-axis, facet, facet_wrap SYMBOL
#legend - stage. y=total count, FACet strain
ck <- ggplot(DMC1.DF[DMC1.DF$stage != "P",], aes(x = stage, y = total.count)) +  facet_wrap(~strain)+ geom_boxplot()+
  stat_compare_means(comparisons = list(c("PWD", "KAZ")))


# annotation table with adjusted pvals and y-position of the labels
anno_df = compare_means(total.count ~ stage, group.by = "strain", data = DMC1.DF[DMC1.DF$stage != "P",]) %>%
 mutate(y_pos = 320)

ck2 <- ggplot(DMC1.DF[DMC1.DF$stage != "P",], aes(x=stage, y=total.count)) + 
  geom_boxplot(position=position_dodge()) + 
  geom_point(aes(color=stage), position=position_jitterdodge()) + 
  facet_wrap(~strain) + 
  ggsignif::geom_signif(
    data=anno_df, 
    aes(xmin=group1, xmax=group2, annotations=p.adj, y_position=y_pos), 
    manual=TRUE
  )
#the plot above adds the arm for (t-test)? comparison of each stage by strain
#this is not quite what I want -- but I'm glad I learned it

#I want a test between 1) high vs low cell stages, 
#I think a better way to display would be
#facet by stage, x is the strain
#compare means across high vs low  (this level isn't defined)

#example from here
#https://github.com/kassambara/ggpubr/issues/65

#  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
#        panel.background = element_blank(), axis.line = element_line(colour = "black"))

DMC1.boxplot
```

Foci of DMC1 were scored from Leptotene and Zygotene spermatocytes of juvenile mice (12-14-18 days). One mouse represents each strain. The main comparison to examine is that between the high (PWD and MSM) and low recombining (WSB, G, KAZ) strains.
Since the total counts are high -- we assume normality and apply t.tests

how similar are G-WSB or PWD-MSM to each other?

ToDo think of ways to add in the compared means for the high and low groups (across cell stage)


## Testing Normality


- check that distributions are close to normal
- check for equal variance.


```{r normality, echo=FALSE}
#histograms for each


#checking equal variance, https://www.statmethods.net/stats/anovaAssumptions.html
#bartlet's test, Levine's test


#bartlett.test(total.count~strain, data=DMC1.DF[DMC1.DF$stage == "L",])#L (don't work with single KAZ L cell)
bartlett.test(total.count~strain, data=DMC1.DF[DMC1.DF$stage == "Z",])#Z

```




From ANOVA analysis strain is significant for Leptotene cells but not Zygotene cells.

```{r DMC1_block1, echo=FALSE, warning=FALSE}
#1. ANOVA, is there a difference in DMC1 counts across strains?
mod.L <- lm(total.count ~ strain, data = DMC1.DF[DMC1.DF$stage == "L",])
mod.Z <- lm(total.count ~ strain, data = DMC1.DF[DMC1.DF$stage == "Z",])

l.anova <- anova(mod.L) #0.0002701 ***
z.anova <- anova(mod.Z) #NS

l.anova
z.anova
```





The two groups of mice will be split into high and low recombining groups. (to assess the patterns of variance across them)

```{r block2_DMC1_t.tests, echo=FALSE, warning=FALSE}
#2. t-tests for the DMC1 counts 
#need to account for sample sizes when comparing t.tests (KAZ and G don't have equal across )
#best case scnario -- more microscoping to get ~20 cells for each stage by strain combo


#L vs Z (all species)
#jj <- t.test(DMC1.male.data$total.count[DMC1.male.data$stage == "Z"], DMC1.male.data$total.count[DMC1.male.data$stage == "L"] )
#p-value = 1.032e-05
#not a valid interpertation


#high vs low group (L)
ttest.L <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
               DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )
# p-value =  0.002145


#high vs low group (Z)
ttest.Z <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
               DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )
# p-value = 0.6628



#dom vs musc (L) (2 subsp vs)
#cheer <- t.test(DMC1.male.data$total.count[(DMC1.male.data$subsp == "Dom") & (DMC1.male.data$stage == "L")], #DMC1.male.data$total.count[(DMC1.male.data$subsp == "Musc") & (DMC1.male.data$stage == "L")] )
# p-value = 0.003016

#dom vs musc (Z)
#up <- t.test(DMC1.male.data$total.count[(DMC1.male.data$subsp == "Dom") & (DMC1.male.data$stage == "Z")], #DMC1.male.data$total.count[(DMC1.male.data$subsp == "Musc") & (DMC1.male.data$stage == "Z")] )
#p-value = 0.09588


#KAZ vs Dom (Z) (only 1 L cell)
#dont <- t.test(DMC1.male.data$total.count[(DMC1.male.data$subsp == "Dom") & (DMC1.male.data$stage == "Z")], 
#      DMC1.male.data$total.count[(DMC1.male.data$strain == "KAZ") & (DMC1.male.data$stage == "Z")] )
#p-value = 0.004148
#Kaz Z are higher than Dom
```

Divide the Musculus strains into high and low recombining groups to characterize the difference in distributions. The p-values from the t.tests of the high vs low groups, indicate that the high recombining are significantly higher for the L cells (p value = `r ttest.L$p.value`) while the Z cells, there is not a significant difference across the high and low groups (value = `r ttest.Z$p.value`).


## Permutations with equal sampling

The number of observations are not quite equal across strains and cell stages, so I ran permutations sampling 25 cells from the high and low recombining groups of cells.


Below are the distributions of p values for permuted t-tests from sampling 25 cells from the pooled high and low recombining groups. Permutations were done for both cell stages. The actual p value is in red. 

The observed p value does not fall outside of the permuted distribution with assumes equal cell sampling.

```{r block3.DMC1.subsampling, echo=FALSE}

#recreate the DMC1 data by subsampling ~9 or lowest cell number from 
# the 2 groups

#make the 2 pools
High.L.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ]
High.Z.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ]

Low.L.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]
Low.Z.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]


#permute drawing / subsampling and performing the t-test
#put the p.value results into a DF!


#permut.SC.cv.diff  <- as.data.frame(
#sample(THE.DF, SAMPLE.SIZE, REPLACE?)

#permuting samples -- and re-licating t-tests

#nrep <- 1000

#permuting samples to replicate the t.test across cell stages from high and low with equal sample sizes.
bunchOpvals.L <- replicate(1000, 
    # t.test
    t.test( sample(High.L.pool, 25, replace = FALSE),
            sample(Low.L.pool, 25, replace = FALSE), var.equal = FALSE )$p.value
          )
bunchOpvals.L <- as.data.frame(bunchOpvals.L)
bunchOpvals.L$stage <- "L"

bunchOpvals.Z <- replicate(1000, 
    # t.test
    t.test( sample(High.Z.pool, 25, replace = FALSE),
            sample(Low.Z.pool, 25, replace = FALSE), var.equal = FALSE)$p.value
          )

bunchOpvals.Z <- as.data.frame(bunchOpvals.Z)
bunchOpvals.Z$stage <- "Z"

#could merge the lists of p-values
replicated.Ttests <- cbind(bunchOpvals.L, bunchOpvals.Z)
#have to melt, the rbinded dataframe to get the cols 'stage' and 'pvale'


#ASSUMPTIONS; normality, equal variance

#plot the lists of pvals in histogram? compare to the observed pvalue to get a sense as of how much sample size shifts the pvalues.

plot.pvals.L <- ggplot(bunchOpvals.L, aes(bunchOpvals.L) )+geom_histogram(bins = 50)+ ggtitle("1000 rep t-tests sample size 25, High vs Low Leptotene")+geom_vline(xintercept=ttest.L$p.value, linetype="dotted", color="red")


plot.pvals.Z <- ggplot(bunchOpvals.Z, aes(bunchOpvals.Z) )+geom_histogram(bins = 50)+ ggtitle("1000 rep t-tests sample size 25, High vs Low Zygotene")+geom_vline(xintercept=ttest.Z$p.value, linetype="dotted", color="red")
plot.pvals.L
plot.pvals.Z


#The t.test( ) function produces a variety of t-tests. Unlike most statistical packages, the default assumes #unequal variance and applies the Welsh df modification.
#n use the var.equal = TRUE
#https://www.statmethods.net/stats/ttest.html
```





> At what stage are DMC1 foci numbers most correlated with MLH1?

Need to pull the strain average table


Leptotene is the max number of DMC1 so this would have more power to capture the variation in MLH1? Look into what baier et al do. 
