---
title: "Effects"
author: "April Peterson"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---

```{r setup, include=FALSE}
#setwd("C:/Users/alpeterson7/Documents/MLH1repo/") #this wd is different than where the file is ... this kinda causes errors
#setwd("C:/Users/April/Desktop/MLH1repo")
#load(file="data/MLH1/MLH1_data_setup.RData")

library(knitr)
library(ggplot2)
library(plyr)
library(lattice)
library(dplyr)
library(raster)
library(lme4)
library(nlme)
library(RLRsim)
library(ggpubr)
library(reshape2)
library(lubridate)#for dates / months

#load main data file

load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_1.5.20.RData")#added batch18

#clean up total SC
load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/total_SC_1.6.20.RData")#total SC data

#load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_12.20.19.RData")#fixing bugs, formating
#load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/total_SC_30.12.19.RData")#total SC data

#load(file="data/MLH1/MLH1_data_setup_11.12.19.RData")
#save.image("~./MLH1repo/data/MLH1/MLH1_data_setup_12.3.19.RData")
```





```{r setupdata, echo=FALSE}
AP_mouse_table_w.Ages <- add_category(AP_mouse_table_w.Ages)
#remove NAs
AP_mouse_table_w.Ages <- AP_mouse_table_w.Ages[!(is.na(AP_mouse_table_w.Ages$Nmice) | AP_mouse_table_w.Ages$Nmice==""), ]

```



# Cell Quality Effects


I am testing the quality effects, since it could be a source of error and potentially biased the predicted gwRR if some mice had bad technical noise.

The the plots below, each point is a single mouse's mean MLH1 count per cell across each quality score bin. 

```{r MLH1.histogram.2, echo=FALSE, include=FALSE}
#thse plots not shown
#remake the pooled MLH1 histgoram plot from above but facet by quality

#PWD male, MSM male
#LEW female, WSB female
mean.MLH1.obs_qual <- ddply(.data=MLH1_data, 
                 .(category, quality),
                 summarize, 
                 mean.MLH1 = mean(adj_nMLH1.foci),
                 var.MLH1 = var(adj_nMLH1.foci)
)
#what's the point of the above table? -- add vlines?

#I think I should do some t-tests across the groups of quality bins to better asess the amount of quality effect

histM_qual.KAZ_male <- ggplot(data=MLH1_data[MLH1_data$category == "KAZ male",],(aes(x=adj_nMLH1.foci))) + xlim(c(16, 38)) + geom_histogram(stat = "bin", binwidth = 1) + facet_wrap(~ quality, ncol = 5)+ggtitle("KAZ male Pooled Cell Histograms by quality")+ geom_vline(xintercept=mean.MLH1.obs_qual$mean.MLH1[mean.MLH1.obs_qual$category == "KAZ male"],color="red")
histM_qual.KAZ_male


histM_qual.PWD_male <- ggplot(data=MLH1_data[MLH1_data$category == "PWD male",],(aes(x=adj_nMLH1.foci))) + xlim(c(16, 38)) + geom_histogram(stat = "bin", binwidth = 1) + facet_wrap(~ quality, ncol = 5)+ggtitle("PWD male Pooled Cell Histograms by quality")+ geom_vline(xintercept=mean.MLH1.obs_qual$mean.MLH1[mean.MLH1.obs_qual$category == "PWD male"],color="red")
histM_qual.PWD_male

histM_qual.MSM_male <- ggplot(data=MLH1_data[MLH1_data$category == "MSM male",],(aes(x=adj_nMLH1.foci))) + xlim(c(16, 38)) + geom_histogram(stat = "bin", binwidth = 1) + facet_wrap(~ quality, ncol = 5)+ggtitle("MSM male Pooled Cell Histograms by quality")+geom_vline(xintercept=mean.MLH1.obs_qual$mean.MLH1[mean.MLH1.obs_qual$category == "MSM male"], color="red")
histM_qual.MSM_male

histM_qual.WSB_female <- ggplot(data=MLH1_data[MLH1_data$category == "WSB female",],(aes(x=adj_nMLH1.foci))) + xlim(c(16, 38)) + geom_histogram(stat = "bin", binwidth = 1) + facet_wrap(~ quality, ncol = 5)+ggtitle("WSB female Pooled Cell Histograms by quality")+ geom_vline(xintercept=mean.MLH1.obs_qual$mean.MLH1[mean.MLH1.obs_qual$category == "WSB female"],color="red")
histM_qual.WSB_female

histM_qual.LEW_female <- ggplot(data=MLH1_data[MLH1_data$category == "LEW female",],(aes(x=adj_nMLH1.foci))) + xlim(c(16, 38)) + geom_histogram(stat = "bin", binwidth = 1) + facet_wrap(~ quality, ncol = 5)+ggtitle("LEW female Pooled Cell Histograms by quality")+ geom_vline(xintercept=mean.MLH1.obs_qual$mean.MLH1[mean.MLH1.obs_qual$category == "LEW female"],color="red")
histM_qual.LEW_female


#plot.pvals.L <- ggplot(bunchOpvals.L, aes(bunchOpvals.L) )+geom_histogram(bins = 50)+ ggtitle("1000 rep t-tests sample size 25, High vs Low Leptotene")+geom_vline(xintercept=ttest.HighLow.L$p.value, linetype="dotted", color="red")


#when I try to add the mean lines, error returned that it can't find the adj_MLH1 ....?
#histM <- histM +  geom_line(data=mean.MLH1.obs, aes(y=mean.MLH1.obs$mean.MLH1))
  
#ad
#  geom_smooth(method = lm, se = FALSE
 # geom_smooth(method = lm, formula = mean.MLH1.obs$mean.MLH1 ~ mean.MLH1.obs$category, se = FALSE)
```


```{r qual.tables, echo=FALSE}

#make a similart plot as the ages
#for each mouse and quality, calq the mean MLH1
mouse.table.mean.MLH1.obs_qual <- ddply(.data=MLH1_data, 
                 .(mouse, quality),
                 summarize, 
                 ncells =  length(unique(Original.Name)),
                 mean.MLH1 = mean(adj_nMLH1.foci),
                 var.MLH1 = var(adj_nMLH1.foci)
)


mouse.table.mean.MLH1.obs_qual <- add_category(mouse.table.mean.MLH1.obs_qual)
mouse.table.mean.MLH1.obs_qual <- add_strain(mouse.table.mean.MLH1.obs_qual)
mouse.table.mean.MLH1.obs_qual <- add_sex(mouse.table.mean.MLH1.obs_qual)
mouse.table.mean.MLH1.obs_qual <- add_subsp(mouse.table.mean.MLH1.obs_qual)

```


```{r qual.plot, echo=FALSE}

mm <- mouse.table.mean.MLH1.obs_qual[mouse.table.mean.MLH1.obs_qual$sex == "male",]
age.effect.plot.mm <- ggplot(mm, aes(y=mean.MLH1, x=quality, color=(strain) ))+geom_jitter(width = 0.16)+ 
  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Male quality and mean CO")+facet_wrap(~subsp)+
  theme(legend.position="none")

age.effect.plot.mm


ff <- mouse.table.mean.MLH1.obs_qual[(mouse.table.mean.MLH1.obs_qual$sex == "female"),]
ff <- ff[ff$strain != "CAST",]

age.effect.plot.ff <- ggplot(ff, aes(y=mean.MLH1, x=quality, color=(strain) ))+geom_jitter(width = 0.16)+ 
  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+
  #are the lines
  ggtitle("Female quality and mean CO")+facet_wrap(~subsp)+theme(legend.position="none")

age.effect.plot.ff

```


These plots show that generally there is a positive relationship between good quality (1) and more MLH1 counts.

I should focus on addressing if there are differences in the relationship of quality-MLH1 counts across sexes.


## Cell Quality effect on SC

Examine image quality or other scores to test how well the python scripts segment the SC area for all the images. 


```{r SC.qual, echo=FALSE}
#use the SC table

#all of skel and bin are not num
MLH1.merge.Skel$skel_size <- as.character(MLH1.merge.Skel$skel_size)
MLH1.merge.Skel$skel_size <- as.numeric(MLH1.merge.Skel$skel_size)

MLH1.merge.Skel$bin_size <- as.character(MLH1.merge.Skel$bin_size)
MLH1.merge.Skel$bin_size <- as.numeric(MLH1.merge.Skel$bin_size)

#mouse.level.skel

#MLH1.merge.Skel, hist of all skel values


#test with removing crop == 'no', MLH1.merge.Skel[MLH1.merge.Skel$REDO.crop == "no",]
mouse.table.mean.skel.obs_qual <- ddply(.data=MLH1.merge.Skel, 
                 .(mouse, quality),
                 summarize, 
                 ncells =  length(unique(Original.Name)),
                 mean.skel = mean(skel_size, na.rm = TRUE),
                 mean.bin = mean(bin_size)
                # var.skel = var(skel_size, na.rm = TRUE) #var not useful, many NA/1 obs
)


mouse.table.mean.skel.obs_qual <- add_category(mouse.table.mean.skel.obs_qual)
mouse.table.mean.skel.obs_qual <- add_strain(mouse.table.mean.skel.obs_qual)
mouse.table.mean.skel.obs_qual <- add_sex(mouse.table.mean.skel.obs_qual)
mouse.table.mean.skel.obs_qual <- add_subsp(mouse.table.mean.skel.obs_qual)

```




```{r SC.hitograms, echo=FALSE, fig.width = 4, fig.height=1}
#try adjusting the size for thi..

skel.histgram <- ggplot(MLH1.merge.Skel, aes(x=skel_size)) +  geom_histogram(bins=50)+ggtitle("width 4, height 1")
skel.histgram
```

```{r more.SC.hist, echo=FALSE}

#length(MLH1.merge.Skel$Random.Name[MLH1.merge.Skel$skel_size > 30000])#67 large SC cells

MLH1.merge.Skel <- add_category(MLH1.merge.Skel)
MLH1.merge.Skel <- add_strain(MLH1.merge.Skel)
MLH1.merge.Skel <- add_sex(MLH1.merge.Skel)
MLH1.merge.Skel <- add_subsp(MLH1.merge.Skel)

MLH1.merge.Skel <- MLH1.merge.Skel[MLH1.merge.Skel$skel_size < 30000,]

skel.histgrams.2.DOM <- ggplot(MLH1.merge.Skel[MLH1.merge.Skel$subsp == "Dom",], aes(x=skel_size)) +  geom_histogram(bins=50)+facet_grid(sex~strain)
skel.histgrams.2.DOM


skel.histgrams.2.MUSC <- ggplot(MLH1.merge.Skel[MLH1.merge.Skel$subsp == "Musc",], aes(x=skel_size)) +  geom_histogram(bins=50)+facet_grid(sex~strain)
skel.histgrams.2.MUSC

```




```{r sc.qual.plots, echo=FALSE}
skel.mm <- mouse.table.mean.skel.obs_qual[mouse.table.mean.skel.obs_qual$sex == "male",]
skel.mm <- skel.mm[skel.mm$mean.skel < 50000,]

skel.qual.effect.plot.mm <- ggplot(skel.mm, aes(y=mean.bin, x=quality, color=(strain) ))+geom_point()+ 
  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Male quality and mean CO")+facet_wrap(~subsp)
skel.qual.effect.plot.mm
```

Each point is a single mouse

```{r G.plot, echo=FALSE}
Bin.qual.effect.plot.mm.G <- ggplot(skel.mm[skel.mm$strain == "G",],
                           aes(y=mean.bin, x=quality, size= ncells, color=strain))+geom_point()+ 
  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Quality and mean SC (bin) ")+facet_wrap(~mouse)

Bin.qual.effect.plot.mm.G


skel.qual.effect.plot.mm.G <- ggplot(skel.mm[skel.mm$strain == "G",],
                           aes(y=mean.skel, x=quality, size= ncells, color=strain))+geom_point()+ 
  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Quality and mean SC (bin) ")+facet_wrap(~mouse)

```

The mouse 22jun15_G_m2 has strange pattern, investigate these outputs

Some of the pattern goes away when redo.crop is set to equal 'no'.

```{r female.skel.plot, echo=FALSE}

skel.ff <- mouse.table.mean.skel.obs_qual[mouse.table.mean.skel.obs_qual$sex == "female",]

skel.qual.effect.plot.ff <- ggplot(skel.ff, aes(y=mean.bin, x=quality, color=(strain) ))+geom_point()+ 
  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Male quality and mean CO")+facet_wrap(~subsp)

skel.qual.effect.plot.ff


skel.qual.effect.plot.ff.MSM <- ggplot(skel.ff[skel.ff$strain == "MSM",], aes(y=mean.bin, x=quality, color=(strain), size=ncells ))+geom_point()+ 
  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Male quality and mean CO")+facet_wrap(~mouse)


skel.qual.effect.plot.ff.G <- ggplot(skel.ff[skel.ff$strain == "G",], aes(y=mean.bin, x=quality, color=(strain), size=ncells ))+geom_point()+ 
  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Male quality and mean CO")+facet_wrap(~mouse)

skel.qual.effect.plot.ff.PWD <- ggplot(skel.ff[skel.ff$strain == "PWD",], aes(y=mean.skel, x=quality, color=(strain), size=ncells ))+geom_point()+ 
  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Male quality and mean CO")+facet_wrap(~mouse)

```

The above plots are good for detecting outliers


# Age Effect (male)

Some papers have documented age effects for male genome-wide recombination rate [@zelazowski2017age].

The initial analysis uses plots. This should be double checked with a general linear model as well.


```{r age.hists, echo=FALSE, eval=FALSE,  warning=FALSE}
source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
meta.data$DOB <- as.Date(meta.data$DOB, format= "%m/%d/%Y")#%Y 2000, %y '18,

#make Mouse meta data work .. rename these
MouseMetaData.wrkin <- meta.data[,c(1,4)]
colnames(MouseMetaData.wrkin) <- c("mouse", "DOB")
MouseMetaData.wrkin <- MouseMetaData.wrkin[!(is.na(MouseMetaData.wrkin$mouse)|MouseMetaData.wrkin$mouse==""),]
#do this for the mouse table instead of the MLH1 table
MLH1.imprv <- merge(MLH1_data, MouseMetaData.wrkin)#don't apply the all parameters
#DOB is date
MLH1.imprv <- add_euth_date(MLH1.imprv)
#good euth date..

#make backup date col #add euth date for calculating age
MLH1.imprv$euth.date2 <- MLH1.imprv$euth.date
MLH1.imprv <- add_age(MLH1.imprv)
#age works

#remove mice that are older than 20 weeks
MLH1.imprv.male <- MLH1.imprv[MLH1.imprv$sex == "male", ]
MLH1.imprv.male <- MLH1.imprv.male[MLH1.imprv.male$age.weeks < 20, ]

age.hist <- ggplot(MLH1.imprv[MLH1.imprv$sex == "male", ], aes(age.weeks))+geom_histogram()+facet_wrap(~category)+ggtitle("Male Age and MLH1 counts by subspecies")
age.hist

```


I made 3 models for 3 groupings of mice, dom, musc.high and musc.low 

$$mouse .mean CO ~=~ age + strain + \varepsilon$$

```{r age.effect.model.setupdata, echo=FALSE}
#run a age effect linear regression!
male.data <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$sex.y == "male",]

#remove all nas

#unorder the factors
class(male.data$strain.x) <- "factor"
class(male.data$subsp.x) <- "factor"

male.data.dom <- male.data[male.data$subsp.y == "Dom",]
male.data.dom <- male.data.dom[!(is.na(male.data.dom$Nmice) | male.data.dom$Nmice==""), ]


#MOLF is a low strain
male.data.musc.low <- male.data[(male.data$strain.x == "KAZ" | male.data$strain.x == "TOM" | male.data$strain.x == "AST" | male.data$strain.x == "MOLF"),]
male.data.musc.low <- male.data.musc.low[!(is.na(male.data.musc.low$Nmice) | male.data.musc.low$Nmice==""), ]


male.data.musc.high <- male.data[(male.data$strain.x == "PWD" | male.data$strain.x == "MSM" | male.data$strain.x == "SKIVE"),]
male.data.musc.high <- male.data.musc.high[!(is.na(male.data.musc.high$Nmice) | male.data.musc.high$Nmice==""), ]


male.data.outgroup <- male.data[(male.data$subsp.y == "Spretus" | male.data$subsp.y == "Spic" | male.data$subsp.y == "Caroli"),]
male.data.outgroup <- male.data.outgroup[!(is.na(male.data.outgroup$Nmice) | male.data.outgroup$Nmice==""), ]

```



```{r age.models, echo=FALSE}

lm.me.dom  <- lm(mean_co ~ age.weeks+strain.x, data = male.data.dom)

lm.me.musc.low  <- lm(mean_co ~ age.weeks+strain.x, data = male.data.musc.low)

lm.me.musc.high  <- lm(mean_co ~ age.weeks+strain.x, data = male.data.musc.high)
#significant differences across strains in musc high

lm.me.OG  <- lm(mean_co ~ age.weeks+strain.x, data = male.data.outgroup)

#summary(lm.me.OG)
as.data.frame( round(  summary(lm.me.dom)$coefficients, 6 ) )

as.data.frame( round(  summary(lm.me.musc.low)$coefficients, 6 ) )

as.data.frame( round(  summary(lm.me.musc.high)$coefficients, 6 ) )
```


ADD euthdate?

Dom, no significant effects for age.  

Musc low, no significant effect for strain or age.  

Musc high, MSM and SKIVE have significant age effects on mean CO count

SKIVE has the greatest spread in ages for single mice (7,16,41 weeks)
There are more SKIVE males which I haven't quantified (at least 3, which are in a good age range)


```{r age.effect.plot, warning=FALSE, echo=FALSE}
source("~./MLH1repo/src/CommonFunc_MLH1repo.R")

#AP_mouse_table_w.Ages <- add_category(AP_mouse_table_w.Ages)
#remove NAs
#AP_mouse_table_w.Ages <- AP_mouse_table_w.Ages[!(is.na(AP_mouse_table_w.Ages$Nmice) | #AP_mouse_table_w.Ages$Nmice==""), ]

#make sure the right dataset 
male.data <- AP_mouse_table_w.Ages[(AP_mouse_table_w.Ages$sex.y == "male") & (AP_mouse_table_w.Ages$category.y != "other"),]
#lots of NA rows... ugh .. remove them
male.data <- male.data[!(is.na(male.data$Nmice) | male.data$Nmice==""), ]

#other removed, no points,
male.data <- male.data[male.data$Ncells > 10,]

age.effect.plot <- ggplot(male.data, aes(y=mean_co, x=age.weeks, color=(strain.y) ))+geom_point()+ ylim(c(19,34))+
  #trying this
  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Male Age and mean CO")+facet_wrap(~subsp.y)
age.effect.plot
#no longer has 'other', #yay I added purtty standard colors
#there is an NA
#todo add line to strain.y

```

I like this plot and find it effective. Can I do a similar set up for other factors (quality, ncells?)

While many slopes are different than 0, I am not convinced there is a clear age effect. Most strains don't have more than 2 ages. I will double check some of these mice and check a general linear model.


# Age Effect (female)

Compare embryonic to neonate (logistic regression). I don't think my embyonic stage estimates are very accurate, so I'll bin all embryos into 1.

```{r female.age, eval=FALSE}
#compile this dataset
```



# Maternal Age effect

Check if there is a maternal age effect for females and males.

```{r mat.age.effect, echo=FALSE, eval=FALSE}
#pause evaluation -- needs debugging
meta.data.file = read.csv("~./MLH1repo/data/DandM_8.30.19.csv", header = TRUE)

meta.data.file <- meta.data.file[!(is.na(meta.data.file$mouse)|meta.data.file$mouse==""),]
#445? -- probably extras at bottom not good

#how many mice have maternal DOB
meta.data.file.w.mat.DOB <- meta.data.file[!(is.na(meta.data.file$maternal_DOB)|meta.data.file$maternal_DOB==""),]
#151

#format this meta.data.file
source("~./MLH1repo/src/CommonFunc_MLH1repo.R")

meta.data.file.w.mat.DOB <- add_mouse(meta.data.file.w.mat.DOB)
meta.data.file.w.mat.DOB <- add_category(meta.data.file.w.mat.DOB)
meta.data.file.w.mat.DOB <- add_strain(meta.data.file.w.mat.DOB)
meta.data.file.w.mat.DOB <- add_sex(meta.data.file.w.mat.DOB)
meta.data.file.w.mat.DOB <- add_subsp(meta.data.file.w.mat.DOB)

#check if date format's been applied before tryung the age functions
#str(meta.data.file.w.mat.DOB)

#DOB and Mat.DOB need to be date formated, make sure that the dates are consistant in the original file
#the dates in my file are not all standardized... year is somethimes 2 - sometimes 4
meta.data.file.w.mat.DOB$DOB <- as.Date(meta.data.file.w.mat.DOB$DOB, format= "%m/%d/%Y")#%Y 2000, %y '18,
meta.data.file.w.mat.DOB$maternal_DOB <-as.Date(meta.data.file.w.mat.DOB$maternal_DOB, format= "%m/%d/%Y")

#str(meta.data.file.w.mat.DOB)

meta.data.file.w.mat.DOB <- add_euth_date(meta.data.file.w.mat.DOB)
meta.data.file.w.mat.DOB <- add_age(meta.data.file.w.mat.DOB)
meta.data.file.w.mat.DOB <- add_mat_age(meta.data.file.w.mat.DOB)

#merge meta data with MLH1 -- rename this
see.whatcha.do <- merge(AP_mouse_table_w.Ages, meta.data.file.w.mat.DOB, by.x = 'mouse', by.y = 'mouse',all = FALSE )
#remove anything without mean_CO
see.whatcha.do <- see.whatcha.do[!(is.na(see.whatcha.do$mean_co)|see.whatcha.do$mean_co==""),]
see.whatcha.do <- see.whatcha.do[!(is.na(see.whatcha.do$mat.age.days)|see.whatcha.do$mat.age.days==""),]
#what did I do with this DF...?

#visualize the data
meta.data.file.w.mat.DOB.female <- meta.data.file.w.mat.DOB[meta.data.file.w.mat.DOB$sex == "female",]
meta.data.file.w.mat.DOB.male <- meta.data.file.w.mat.DOB[meta.data.file.w.mat.DOB$sex == "male",]


#hist of data
mat.age.hist.female <- ggplot(meta.data.file.w.mat.DOB.female, aes(mat.age.days))+geom_histogram()+facet_wrap(~category)+ggtitle("Maternal ages of females")
mat.age.hist.female

mat.age.hist.male <- ggplot(meta.data.file.w.mat.DOB.male, aes(mat.age.days))+geom_histogram()+facet_wrap(~category)+ggtitle("Maternal ages of males")
mat.age.hist.male


#remove mice that might have low quality
see.whatcha.do.HQ <- see.whatcha.do[see.whatcha.do$Ncells > 14,]

#age by mean_co (these aren't working)
mat.age.effect.plot.female <- ggplot(see.whatcha.do[see.whatcha.do$sex == "female",], aes(y=mean_co, x=mat.age.days, color=(strain.y) ))+geom_point()+ ylim(c(19,34))+
  #trying this
  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Female Materinal Age and Mean CO (offsping)")+facet_wrap(~subsp.y)
mat.age.effect.plot.female

mat.age.effect.plot.male <- ggplot(see.whatcha.do.HQ[see.whatcha.do.HQ$sex == "male",], aes(y=mean_co, x=mat.age.days, color=(strain.y) ))+geom_point()+ ylim(c(19,34))+
  #trying this
  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Maternal Age and mean CO, male offspring")+facet_wrap(~subsp.y)

#
ALLmat.age.effect.plot.male <- ggplot(see.whatcha.do[see.whatcha.do$sex == "male",], aes(y=mean_co, x=mat.age.days, color=(strain.y) ))+geom_point()+ ylim(c(19,34))+
  #trying this
  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Maternal Age and mean CO, male offspring")
+facet_wrap(~subsp.y)

#is there a correlation with proband age and maternal age?
#there is a big lack of maternal age data...
#some of the values reference the original breeding pair..
#from notes on original BPs(males), assigning maternal ages

```




### Litter effect

Compare female spreads made on the same day, but from different mothers/litters.
```{r litter.effect, eval=FALSE}
#compile this data set

```


# Dissection Effect




```{r dissection.effect.data.setup, echo=FALSE}

#AP_mouse_table_w.Ages$euth.date

mouse.table.female <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$sex.x == "female",]

mouse.table.female <- mouse.table.female[!(is.na(mouse.table.female$mouse) | mouse.table.female$mouse==""), ]

mouse.table.female <- mouse.table.female[!(is.na(mouse.table.female$euth.date) | mouse.table.female$euth.date==""), ]


#add euth dates to MLH1_data
test.this <- MLH1_data[ MLH1_data$mouse %in% mouse.table.female$mouse, ]

mlh1.euth.date.female <- merge(mouse.table.female, MLH1_data, by = "mouse")

```


```{r plot.dissection, echo=FALSE}

male.data$euth.date <- as.Date(male.data$euth.date, "%m/%d/%Y")

male.euth.date <- ggplot(male.data, aes(y=mean_co, x=euth.date, color=(strain.y) ))+geom_point()+ ylim(c(19,34))+
  #trying this
  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Male Age and mean CO")+facet_wrap(~subsp.y)
male.euth.date


#look at musc
musc.male.euth.date <- ggplot(male.data[male.data$subsp.x == "Musc",], aes(y=mean_co, x=euth.date, color=(age.weeks) ))+geom_point()+ ylim(c(19,34))+
  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Musc Male Age and mean CO")+facet_wrap(~strain.x)
musc.male.euth.date

```



```{r dissection.female, echo=FALSE}

female.data <- AP_mouse_table_w.Ages[(AP_mouse_table_w.Ages$sex.y == "female") & (AP_mouse_table_w.Ages$category.y != "other"),]
female.data$euth.date <- as.Date(female.data$euth.date, "%m/%d/%Y")

female.euth.date.plot <- ggplot(female.data[female.data$subsp.x == "Musc",], aes(y=mean_co, x=euth.date, color=(strain.y) ))+geom_point()+
  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Male Age and mean CO")+facet_wrap(~subsp.y)

female.euth.date.plot


DOM.female.euth.date.plot <- ggplot(female.data[female.data$subsp.x == "Dom",], aes(y=mean_co, x=euth.date, color=(strain.y) ))+geom_point()+
  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Female Age and mean CO")+facet_wrap(~subsp.y)

DOM.female.euth.date.plot


#female / male, sex.x
#mouse.table.male <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$sex.x == "male",]

test.this <- MLH1_data[ MLH1_data$mouse %in% mouse.table.female$mouse, ]

mlh1.euth.date.female <- merge(mouse.table.female, MLH1_data, by = "mouse")

```


## Total SC

Testing the effects of mean total SC per mouse with when they were dissected.

follow up to peromyscus findings, Where summer caught mice had lower SC area per cell, compared to winter and fall caught. But there were few data points.


```{r total.sc, echo=FALSE}
#check out effects of month dissected on SC
#load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_12.20.19.RData")#fixing bugs, formating
#from total SC

#merge
ll <- merge(mouse.level.skel, AP_mouse_table_w.Ages, by = "mouse")  #no
ll$euth.date.2 <- as.Date(ll$euth.date, "%m/%d/%Y")
#ncells is from the full MLH1 DF
#Ncells is from the cleaned, HQ DF

ll <- ll[!(is.na(ll$mean.skel) | ll$mean.skel==""), ]
ll <- ll[!(is.na(ll$sex.y) | ll$sex.y==""), ]

ll <- subset(ll, select = -c(strain.x,strain.y, category.x, category.y, sex.x, sex.y, subsp.x, subsp.y, mouse.1, nmice) )

#ADD month number
ll$month.2 <- month(ll$euth.date.2)

ll.plot <- ggplot(ll, aes(y=mean.skel, x=euth.date.2, color=(strain) ))+geom_point()+
 scale_color_manual(values=colors_of_strains)+
  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("")+facet_wrap(~sex)

ll.month.plot <- ggplot(ll, aes(y=mean.skel, x=month.2, color=(strain) ))+geom_point()+
  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+
  ggtitle("")+facet_wrap(~sex)

ll.month.plot

```


```{r total.sc.all.cells, echo=FALSE, include=FALSE}
#try ploting all cells

#uu <- merge(MLH1.merge.Skel,AP_mouse_table_w.Ages,by="mouse"  )

#uu$euth.date.2 <- as.Date(uu$euth.date, "%m/%d/%Y")
#uu$month.2 <- month(uu$euth.date.2)

#uu <- uu[!(is.na(uu$skel_size) | uu$skel_size==""), ]
#uu <- uu[!(is.na(uu$sex.x) | uu$sex.x==""), ]

#uu.month.plot.DOM <- ggplot(uu[uu$subsp.x == "Dom",], aes(y=skel_size, x=month.2, color=(mouse) ))+
#  geom_jitter()+  theme(legend.position="none")+
#  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("")+
#  facet_wrap(~sex.x)

```


So, there doesn't seem to be an effect of month dissected on SC area across mice. This might be the start of testing for 'technical effects / humidity effects'. BUT, I don't really have month by month dissections of mice of the same age, ect.

> *month dissected*

For mice strains that have a similar MLH1 level (control for strain MLH1 effect), check if a month effect for MLH1 count (as seen in Peromyscus).



# Spread and Stain Effects




> *staining batches*

Compare observations from the same mouse, but different MLH1 stains.  First need to identify data that can be used.

```{r stain.effects.SC, eval=FALSE}
#/src/TotalSC.R
#compile this dataset

```


# Room effect

```{r ROOM.effect, eval=FALSE}

#COMPILE THIS datasets
#3 different rooms used to house mice. I should have some data on when the switch was. This would be limited to, WSB, LEW and PWD


```

