---
title: "Power Calculation Report"
author: "April Peterson"
date: "June 12, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Power Calculations

Main questions to address:
 
 > *How many cells per mouse should be sampled?*
 
 > *How many mice per strain should be sampled?*

Outline discription of the loop for comparing sample size effects

i) take ~100 independant samples of total images across different sample sizes (10%, 25%, 50%, 75%, 90%)
ii) then run t.tests across those samples, and report the p value 
iii) make a table of the p values for these sample permutations
iv) plot the p values or mean p values across sample sizes

Currently this section is under construction.


```{r, include=FALSE}
#passed_mice_data <- MLH1_data[MLH1_data$mouse %in% passed_mice, ]
#non_passed_mice_data <- MLH1_data[MLH1_data$mouse %in% not_passed, ]

#make a column to mark if observation comes from passed or not passed mouse
#turn this into boxplot and sort the category order, otherwise it's hard to any patterns
#sc <- qplot(mouse, nMLH1.foci,  colour = category, data = non_passed_mice_data, main="s", xlab="s", ylab="es")
#sc + theme(legend.position="none")

#sc <- qplot(mouse, nMLH1.foci,  colour = category, data = passed_mice_data, main="s", xlab="s", ylab="es")
#sc + theme(legend.position="none")

# not working not sure of the purpose
#count =1
#for(i in MLH1_data$file.name){
#  templist= strsplit(i, split="_")[[1]]
#  if(MLH1_data$mouse %in% passed_mice) 
#  MLH1_data$mouse[count] <- c
#  count= count +1
#}
```