---
title: "MS_Tables"
author: "April Peterson"
date: "6/16/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(xlsx)
library(knitr)
library(kableExtra)
#read in excel plates
Table1 = read.xlsx2("~./MLH1repo/doc/MainTables.xlsx", header=TRUE, 1)

Table2 = read.xlsx2("~./MLH1repo/doc/MainTables.xlsx", header=TRUE, 2)

Table3 = read.xlsx2("~./MLH1repo/doc/MainTables.xlsx", header=TRUE, 3)

Table4 = read.xlsx2("~./MLH1repo/doc/MainTables.xlsx", header=TRUE, 4)

#Supplemental Table
```

```{r table1, echo=FALSE}

#change the headers

colnames(Table1) <- c("Species","Official name","abbreviation for this paper","Geographic Origin","Source")
print.tex <- kable(Table1, "latex")
```

```{r tab2, echo=FALSE}
print.tex2 <- kable(Table2, "latex")
```


```{r table3, echo=FALSE}

#arg tha numbers are factors!!!!
Table3$Number.of.Mice <- as.character(Table3$Number.of.Mice)
Table3$Number.of..Cells <- as.character(Table3$Number.of..Cells)
Table3$Mean.MLH1.Count <- as.character(Table3$Mean.MLH1.Count)
Table3$se <- as.character(Table3$se)
Table3$cV <- as.character(Table3$cV)
Table3$var <- as.character(Table3$var)

Table3$Number.of.Mice <- as.numeric(Table3$Number.of.Mice)
Table3$Number.of..Cells <- as.numeric(Table3$Number.of..Cells)
Table3$Mean.MLH1.Count <- as.numeric(Table3$Mean.MLH1.Count)
Table3$se <- as.numeric(Table3$se)
Table3$cV <- as.numeric(Table3$cV)
Table3$var <- as.numeric(Table3$var)


#MLH1 summary stats
print.tex3 <- kable(Table3, "latex",booktabs = T, digits = 2)

write.table(kable(Table3, "latex",booktabs = T, digits = 2), "~./MLH1repo/Table3.mean.table.tex", sep=",",
            row.names = FALSE)

```



```{r table4, echo=FALSE}
#DMC1 table

colnames(Table4) <- c("Recombination Group","Strain",
                      "Mean DMC1 Foci per Cell","MLH1:DMC1   ratio","MLH1:DMC1   ratio",
                     "Mean DMC1 Foci per Cell","MLH1:DMC1   ratio","MLH1:DMC1   ratio")

#DMC1
print.tex4 <- kable(Table4, "latex", booktabs = T) %>%
kable_styling() %>%
add_header_above(c(" " = 2, "Early Zygotene" = 3, "Late Zygotene" = 3) )

write.table(print.tex4, "~./MLH1repo/DMC1.mean.table.tex")

save_kable(print.tex4, "~./MLH1repo/kable.DMC1.mean.table.html", keep_tex = TRUE)


#kable(dt, "latex", booktabs = T) %>%
#kable_styling() %>%
#add_header_above(c(" " = 1, "Group 1" = 2, "Group 2" = 2, "Group 3" = 2))

#write the kable file


```




\begin{tabular}{S l l l r}
\toprule
{Speed (mph)} & Driver          & Car                        & Engine    & Date     \\
\midrule
407.447     & Craig Breedlove & Spirit of America          & GE J47    & 8/5/63   \\
413.199     & Tom Green       & Wingfoot Express           & WE J46    & 10/2/64  \\
434.22      & Art Arfons      & Green Monster              & GE J79    & 10/5/64  \\
468.719     & Craig Breedlove & Spirit of America          & GE J79    & 10/13/64 \\
526.277     & Craig Breedlove & Spirit of America          & GE J79    & 10/15/65 \\
536.712     & Art Arfons      & Green Monster              & GE J79    & 10/27/65 \\
555.127     & Craig Breedlove & Spirit of America, Sonic 1 & GE J79    & 11/2/65  \\
576.553     & Art Arfons      & Green Monster              & GE J79    & 11/7/65  \\
600.601     & Craig Breedlove & Spirit of America, Sonic 1 & GE J79    & 11/15/65 \\
622.407     & Gary Gabelich   & Blue Flame                 & Rocket    & 10/23/70 \\
633.468     & Richard Noble   & Thrust 2                   & RR RG 146 & 10/4/83  \\
763.035     & Andy Green      & Thrust SSC                 & RR Spey   & 10/15/97\\
\bottomrule
\end{tabular}