---
title: "Chapter3_ideas"
author: "April Peterson"
date: "November 26, 2019"
output: html_document
---

```{r setup, include=FALSE}
#load biv data
#load MLH1 data
library(dplyr)
library(plyr)
library(ggplot2)

```


Take some mouse level distributions -- pull cells from top and bottom 10%?
(gather whole bivalent level data,)

-start making comparisons for the 2 pools
  how much technical noise is there? (staining differences?/ quality score?
  
Mice with many cells and BivData
17mar16_G_f4 - BivData
17mar16_G_f1


```{r choosing.mice, echo=FALSE, eval=FALSE}
#Choose mice with many cells

#G_f, G_m
#SKIVE_f, SKIVE_m #have BivData

#remove all the quality 5

#make plot/histogram of the MLH1 count distribution
counts <- MLH1_data$nMLH1.foci[MLH1_data$mouse == "31may19_SKIVE_f6"]

mouse.DF <- MLH1_data[MLH1_data$mouse == "31may19_SKIVE_f6",]
mouse.DF <- mouse.DF[mouse.DF$quality < 5,]
high.bin.DF <- mouse.DF[mouse.DF$nMLH1.foci>=quantile(counts, 0.9, na.rm=TRUE), ]
low.bin.DF <- mouse.DF[mouse.DF$nMLH1.foci<=quantile(counts, 0.1, na.rm=TRUE), ]

ordered.MLH1 <- mouse.DF[with(mouse.DF,order(nMLH1.foci)),]#this returns ordered
topsy <- ordered.MLH1[1:5,]
bottomzy <- ordered.MLH1[(length(ordered.MLH1$nMLH1.foci)-5):length(ordered.MLH1$nMLH1.foci),]
#now have top and bottom 5 cell observations 

#merge
top.n.bottom.bin <- rbind(topsy,bottomzy)

#write 
write.table(top.n.bottom.bin, "~./MLH1repo/curated.bins.csv", sep=",", row.names = FALSE)


shiz <- ggplot(data = MLH1_data[MLH1_data$mouse == "31may19_SKIVE_f6",],
       aes(x=nMLH1.foci))+geom_histogram(stat = "bin", binwidth = 1)+ggtitle("31may19_SKIVE_f6 Cell counts")+
geom_vline(xintercept = c(min(high.bin), max(low.bin)), color= "red")
#highest of the low bin
#lowest value of the high bin
shiz
#get the filenames for this groups of cells

#10mar15_PWD_m1#10mar15_PWD_m2 -- not found in BivData, re-running all of the PWD male data
#28feb15_PWD_f2, in PWD--female .xlsx file

#31may19_SKIVE_m2
male.one.DF <- MLH1_data[MLH1_data$mouse == "28feb15_PWD_f2",]
male.one.DF <- male.one.DF[male.one.DF$quality < 5,]

ordered.MLH1.male <- male.one.DF[with(male.one.DF,order(nMLH1.foci)),]#this returns ordered
male.high.bin.DF <- ordered.MLH1.male[1:5,]
male.low.bin.DF <- ordered.MLH1.male[(length(ordered.MLH1.male$nMLH1.foci)-5):length(ordered.MLH1.male$nMLH1.foci),]
#now have top and bottom 5 cell observations 

#merge
male.top.n.bottom.bin <- rbind(male.high.bin.DF, male.low.bin.DF)
#write 
write.table(male.top.n.bottom.bin, "~./MLH1repo/female.PWD.curated.bins.csv", sep=",", row.names = FALSE)


shiz.too <- ggplot(data = MLH1_data[MLH1_data$mouse == "31may19_SKIVE_m2",],
       aes(x=nMLH1.foci))+geom_histogram(stat = "bin", binwidth = 1)+ggtitle("31may19_SKIVE_m2 Cell counts")+
geom_vline(xintercept = c(min(male.high.bin.DF$nMLH1.foci), max(male.low.bin.DF$nMLH1.foci)), color= "red")
#highest of the low bin
#lowest value of the high bin
shiz.too


```


the plot above shows an example plot of --- where the pools are from, currently these are 10 quantille and 90 quantile. setting threshold by percentile doesn't always give enough cells. 
Should I be comparing cells with the same number of COs?



```{r getting.cell.names, echo=FALSE,eval=FALSE}

#load in the cell dataset
#28feb15_PWD_f2  from Curated_BivData
#for a mouse, pull the 10 highest rec cells

PWD.female_hand.measure = read.csv("~./MLH1repo/data/BivData/PWD_female_hand_measure.csv", header = TRUE, strip.white = TRUE)
#"fileName"                "boxNumber"               "blobjectClass"           "chromosomeLength"        "centromere_ABS_Position" "Foci1"                   "Foci2"                  
#"Foci3"                   "curation.notes"  
female.mouse_bivData <- Curated_BivData[Curated_BivData$mouse == "28feb15_PWD_f2",]

write.table(female.mouse_bivData, "~./MLH1repo/28feb15_PWD_f2.csv", sep=",", row.names = FALSE)

PWD.female_hand.measure = read.csv("~./MLH1repo/28feb15_PWD_f2.csv", header = TRUE, strip.white = TRUE)
PWD.female_hand.measure$Obj.ID <- paste(PWD.female_hand.measure$fileName, PWD.female_hand.measure$boxNumber, sep = "_")

#merge with hand data
#mergy <- merge(PWD.female_hand.measure, female.mouse_bivData)
#PWD.female_hand.measure$hand.foci.count <- as.numeric(PWD.female_hand.measure$hand.foci.count)


cells.BivData_table_28feb15_PWD_f2 <- ddply(.data=PWD.female_hand.measure, 
                            .(fileName),
                            summarize, 
                            ncells = length(unique(fileName)),
                            nbivs = length(unique(Obj.ID)),
                            n.foci = sum(hand.foci.count),
                            total.SC = sum(chromosomeLength),
                            mean.SC = mean(chromosomeLength)
                            
)
#add hand data
ll <- cells.BivData_table_28feb15_PWD_f2[cells.BivData_table_28feb15_PWD_f2$nbivs > 18,]

#I have 3 cells with full 20

```


Start discribing the Bivalent patterns for the PWD female patterns


```{r mini.biv, echo=FALSE,eval=FALSE}
#make mini biv plots

#PWD.female
#PWD.female_hand.measure

cell.1 <- PWD.female_hand.measure[PWD.female_hand.measure$fileName == "13mar15_28feb15_PWD_f2_sp1_10_rev",]
cell.2 <- PWD.female_hand.measure[PWD.female_hand.measure$fileName == "13mar15_28feb15_PWD_f2_sp1_38_rev",]


two_groups <- transform(cell.1, category2 = factor(paste(fileName, chromosomeLength)))
two_groups <- transform(two_groups, category2 = reorder(category2, rank(chromosomeLength)))

#this plot function takes the new df, and plots cat2(x) by SC length (y). cat2 has been ordered by the ranked length
#of SC.
MiniBiv_plot1 <-  ggplot(two_groups, aes(category2, chromosomeLength)) +
    geom_bar(aes(fill= as.factor(hand.foci.count)), stat = "identity") +
  #  geom_vline(xintercept = c(male_IFD, 1) ) +  #
 #geom_hline(yintercept = c(male_IFD) ) +
  
  scale_fill_manual(values=c("red", "red", "red")) +   #0CO;"#FFCC66"
  #how to add points for 2foci
  geom_point(aes(x=two_groups$category2, y=two_groups$Foci1),colour = "green",size = 3 ) +   
  #foci.1.position, two_groups$distal.foci.position)
  geom_point(aes(x=two_groups$category2, y=two_groups$Foci2),colour = "green",size = 3 ) +
  
#add centromeres...  
  geom_point(aes(x=two_groups$category2, y=0), colour = "purple4",size =5)+


  facet_wrap( ~ fileName, scales = "free_x", ncol = 5) +
 # scale_x_discrete(labels=two_groups$strain, breaks=two_groups$category2) +
  ylim(c(0,200))+
     theme(axis.line=element_blank(),axis.text.x=element_blank(),
        axis.ticks=element_blank())  #for removeing x labels

#MiniBiv_plot1
#MiniBiv_plot2

#save the files 

```



```{r biscribe.Biv.Pattern, echo=FALSE}
#start discribing the basic 

#start merging curated bivData with handmeasured data


#BivData - Curated
```

# Whole Cell Measures

excel sheet Whole_cell_measures.xlsx has 840 bivalent measures.


```{r whole.cell.data, echo=FALSE}

#read in csv of whole cell hand measures
#setwd("C:/Users/alpeterson7/Documents/MLH1repo/")
#setwd("C:/Users/April/Desktop/MLH1repo")
#load(file="data/MLH1/MLH1_data_setup.RData")

#load main data file
whole.cell.data <- read.csv("~./MLH1repo/whole_cell_measures.csv")

source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
#add sex since there is a difference in sample size
whole.cell.data <- add_mouse(whole.cell.data)
whole.cell.data <- add_sex(whole.cell.data)

whole.cell.data <- whole.cell.data[!(is.na(whole.cell.data$fileName) | whole.cell.data$fileName==""), ]

whole.cell.data.XY <- whole.cell.data
whole.cell.data <- whole.cell.data[(whole.cell.data$hand.foci.count!="XY"), ]

#make sure foci.count is number
whole.cell.data$hand.foci.count <- as.character(whole.cell.data$hand.foci.count)
whole.cell.data$hand.foci.count <- as.numeric(whole.cell.data$hand.foci.count)

#find cells with the wrong number

biv.count_table <-  ddply(.data=whole.cell.data,
                 .(fileName),
                 summarize,
                 nbiv = length(SC.length)
)
#all have 20 or 19 (after removing XY)

#ADD THE RANK
for(musy in as.character(unique(whole.cell.data$fileName))){
  print(musy)
  
  #order the chrm lengths -- this is order
  chrm.order = order(whole.cell.data$SC.length[whole.cell.data$fileName == musy])
  whole.cell.data$rank[whole.cell.data$fileName == musy] <- chrm.order #19 to 20
}
#I think I saved the Excel with sorted SC lengths
#Yay I printed the rank in this

```




Try doing the permutations

```{r permutations.sc, echo=FALSE}

male.whole.cell <- whole.cell.data[whole.cell.data$sex == "male", ]#359
#ncells
length(unique(male.whole.cell$fileName) ) #24 cells


female.whole.cell <- whole.cell.data[whole.cell.data$sex == "female", ]#456
length(unique(female.whole.cell$fileName) ) #18 cells

male.biv.count_table <-  ddply(.data=male.whole.cell,
                 .(fileName),
                 summarize,
                 nbiv = length(SC.length),
                 mean_co = mean(hand.foci.count),
                 mean_sc = mean(SC.length),
                   var_co = var(hand.foci.count),
                   var_sc = var(SC.length)
)


#not sure how to make the initial DF
#replicate 24 rows with 19 cols
iterations = 24
variables = 20
DF.permutations.sc <- matrix(ncol=variables, nrow=iterations)
colnames(DF.permutations.sc) <- c("1", "rank1", "rank2", "rank3", "rank4", "rank5",
                               "rank6","rank7","rank8","rank9","rank10","rank11",
                               "rank12","rank13","rank14","rank15","rank16",
                               "rank17","rank18","rank19" )

#ROW IS FIRST
for(row.num in 1:24){
  #print(c(row.num, "rowis") )
  for(col.num in 2:ncol(DF.permutations.sc) ){
  rand.rank.sc <- sample(male.whole.cell$SC.length[male.whole.cell$rank == col.num-1], 1, replace = FALSE)
  #print(SCll)   
  DF.permutations.sc[row.num,col.num] <- rand.rank.sc
   }
  }#yes this finally works!!!!
DF.permutations.sc <- as.data.frame(DF.permutations.sc)


#pair with real within cell SC variance
for( o in 1:nrow(DF.permutations.sc) ) {
  DF.permutations.sc$var_sc[o] <- var(as.numeric(DF.permutations.sc[o,2:20]))
  }



#permutations without accounting for rank
its = 24
varis = 20
DF.permutations.no_rank <- matrix(ncol=varis, nrow=its)
#colnames(DF.permutations.sc) <- c("1", "rank1", "rank2", "rank3", "rank4", "rank5",
#                               "rank6","rank7","rank8","rank9","rank10","rank11",
#                               "rank12","rank13","rank14","rank15","rank16",
#                               "rank17","rank18","rank19" )

for(row.num in 1:24){
  #print(c(row.num, "rowis") )
  for(col.num in 2:ncol(DF.permutations.no_rank) ){
  rand.sc <- sample(male.whole.cell$SC.length[male.whole.cell$rank == col.num-1], 1, replace = FALSE)
  #print(SCll)   
  DF.permutations.no_rank[row.num,col.num] <- rand.sc
   }
  }#yes this finally works!!!!
DF.permutations.no_rank <- as.data.frame(DF.permutations.no_rank)

for( o in 1:nrow(DF.permutations.sc) ) {
  DF.permutations.no_rank$var_rand.sc[o] <- var(as.numeric(DF.permutations.no_rank[o,2:20]))
  }

#merge with real
sc.var.for.plot <- matrix(ncol=3, nrow=72)
sc.var.for.plot <- as.data.frame(sc.var.for.plot)
colnames(sc.var.for.plot) <- c("cell.id","var","type")

sc.var.for.plot$var[1:24] <- DF.permutations.sc$var_sc
sc.var.for.plot$type[1:24] <- "sim"

sc.var.for.plot$var[25:48] <- male.biv.count_table$var_sc
sc.var.for.plot$type[25:48] <- "real"

sc.var.for.plot$var[49:72] <- DF.permutations.no_rank$var_rand.sc
sc.var.for.plot$type[49:72] <- "sim_no_rank"


#plot the forms
#plot within cell variance
whole.cell.var_SC.plot  <- ggplot(sc.var.for.plot, aes(y=var, x=type))+geom_jitter()+ggtitle("within cell sc variance")
whole.cell.var_SC.plot
```



```{r per.co.count, echo=FALSE}

iterations = 24
variables = 20
DF.permutations_CO.count <- matrix(ncol=variables, nrow=iterations)
colnames(DF.permutations_CO.count) <- c("1", "rank1", "rank2", "rank3", "rank4", "rank5",
                               "rank6","rank7","rank8","rank9","rank10","rank11",
                               "rank12","rank13","rank14","rank15","rank16",
                               "rank17","rank18","rank19" )

for(row.num in 1:24){
  
  #print(c(row.num, "rowis") )
  for(col.num in 2:ncol(DF.permutations_CO.count) ){
  
  rand.rank.co <- sample(male.whole.cell$hand.foci.count[male.whole.cell$rank == col.num-1], 1, replace = FALSE)
  #print(SCll)   
  DF.permutations_CO.count[row.num,col.num] <- rand.rank.co
   }
  }
DF.permutations_CO.count <- as.data.frame(DF.permutations_CO.count)


#no rank - iterate / permute the sampling X times
its = 24
varis = 20
DF.permutations.CO.no_rank <- matrix(ncol=varis, nrow=its)

#make LIST of X DATAFRAMES

for(row.num in 1:24){
  #print(c(row.num, "rowis") )
  for(col.num in 2:ncol(DF.permutations.CO.no_rank) ){
  rand.co <- sample(male.whole.cell$hand.foci.count, 1, replace = FALSE)
  #print(SCll)   
  DF.permutations.CO.no_rank[row.num,col.num] <- rand.co
   }
  }#yes this finally works!!!!
DF.permutations.CO.no_rank <- as.data.frame(DF.permutations.CO.no_rank)

#
for( o in 1:nrow(DF.permutations.CO.no_rank) ) {
  DF.permutations.CO.no_rank$var_rand.co[o] <- var(as.numeric(DF.permutations.CO.no_rank[o,2:20]))
  }





#merge with real
co.var.for.plot <- matrix(ncol=3, nrow=72)
co.var.for.plot <- as.data.frame(co.var.for.plot)
colnames(co.var.for.plot) <- c("cell.id","var","type")

co.var.for.plot$var[1:24] <- DF.permutations_CO.count$var_CO.count
co.var.for.plot$type[1:24] <- "sim"

co.var.for.plot$var[25:48] <- male.biv.count_table$var_co
co.var.for.plot$type[25:48] <- "real"

co.var.for.plot$var[49:72] <- DF.permutations.CO.no_rank$var_rand.co
co.var.for.plot$type[49:72] <- "sim_no_rank"


#plot
whole.cell.var_CO.plot  <- ggplot(co.var.for.plot, aes(y=var, x=type))+geom_jitter()+ggtitle("within cell CO variance")
whole.cell.var_CO.plot

```




```{r permuted.v.real, echo=FALSE}
#evaluating the permuted 
library(plyr)

for( o in 1:nrow(DF.permutations_CO.count) ) {
  DF.permutations_CO.count$var_CO.count[o] <- var(as.numeric(DF.permutations_CO.count[o,2:20]))
  }

#two data frames of within cell var
#cell.id, var, type
var.for.plot <- matrix(ncol=3, nrow=48)
var.for.plot <- as.data.frame(var.for.plot)
colnames(var.for.plot) <- c("cell.id","var","type")

var.for.plot$var[1:24] <- DF.permutations_CO.count$var_CO.count
var.for.plot$type[1:24] <- "sim"

var.for.plot$var[24:48] <- male.biv.count_table$var_co
var.for.plot$type[24:48] <- "real"


#my permuted DFs are in a strange str

#plot within cell variance
whole.cell.var.plot  <- ggplot(var.for.plot, aes(y=var, x=type))+geom_jitter()+ggtitle("within cell variance between types")

whole.cell.var.plot
```



```{r junk, eval=FALSE}
DF.permutations <- data.frame( nuc.sim = seq(1:24), rank1 = replicate(24, ""),rank2= replicate(24, ""),
                       rank3= replicate(24, ""),     rank4= replicate(24, ""),
                       rank5= replicate(24, ""),   rank6= replicate(24, ""),
                       rank7= replicate(24, ""), rank8= replicate(24, ""),
                       rank9= replicate(24, ""), rank10= replicate(24, ""),
                       rank11= replicate(24, ""),     rank12= replicate(24, ""),
                       rank13= replicate(24, ""),     rank14= replicate(24, ""),
                       rank15= replicate(24, ""),     rank16= replicate(24, ""),
                       rank17= replicate(24, ""),rank18= replicate(24, ""),rank19= replicate(24, "")
                       )
```

